### 2024-02-29 aboutme.qmd
update website

---
title: "Jackson Strand"
about:
  id: me
  template: trestles
  links:
    - icon: linkedin
      text: LinkedIn
      href: https://linkedin.com/in/jacksonstrand
    - icon: instagram
      text: Instagram
      href: https://www.instagram.com/jstrand894/
    - icon: github
      text: GitHub
      href: https://github.com/jstrand894 
  image: img/aboutme/deadhorse.jpeg

sidebar: false
---

:::{#me}

I am currently a Master's graduate student at Montana State University studying entomology. I live in Bozeman, MT with my wife Grace.

[My CV](Aboutme/jackson_strand_updated_cv.pdf)

![Me and Grace at Glacier National Park, 2023](img/aboutme/iceberg.JPG)

:::