### 2024-02-29 bugs.qmd

---
title: "bug pictures"
filters:
  - lightbox
lightbox: auto
---

![Bug pic 1](img/bugs/IMG_1664.JPG){group="my-gallery"}
![Bug pic 2](img/bugs/IMG_1678.JPG){group="my-gallery"}

<figure>
  <img src="img/bugs/IMG_1664.JPG" alt="Montana State University" 
  width = 200" height = "200">
</figure>