### 2024-02-29 deadhorse.qmd

---
title: "Deadhorse Ultra 50 Miler"
subtitle: "November 18, 2023"
format: html
filters:
   - lightbox
lightbox: auto
editor_options: 
  chunk_output_type: console
---

::: {style="display: grid;grid-template-columns: repeat(auto-fill, minmax(200px, 1fr));grid-gap: 1em;"}
```{r get images, echo = F, results = 'asis'}
image_files <- list.files("img/deadhorse", pattern = ".JPG|.png|.gif|.jpg|.jpeg", full.names = TRUE)

for (file in image_files) {
  cat("![](",file, ")\n  \n")
}
```
:::