### 2024-02-29 field_report.qmd

---
title: "MT Field Site Analysis Report"
author: "Jackson Strand"
date: "`r Sys.Date()`"
output: 
  html_document:
    self_contained: yes
    include:
    keep_md: true
    theme: cosmo
    toc: true
    toc_float: true
    toc_title: "Table of Contents"
    toc_collapsed: false
    toc_depth: 4
    # number_sections: true
editor_options: 
  chunk_output_type: console
---
```{r library, echo = F, message = F, warning = F}
library(tidyverse)
library(readxl)
library(ggplot2)
library(kableExtra)
library(plotly)
```
```{r column.order, echo = F}
column.order <- c("distance", "collection", "p.inf", 
                  "p.cut", "p.par", "poten")

column.order2 <- c("location", "distance", "collection", "p.inf", 
                  "p.cut", "p.par", "poten")

round_based_on_size <- function(x) {
  threshold <- 0.01
  rounded_values <- ifelse(abs(x) > threshold, round(x, 2), round(x, 3))
  return(rounded_values)
}
```

# POST FARM

## Methods

### Plant Selection and Cages

Assessment of C. cinctus infestation and mortality within B. inermis were assessed using a 34 x 60 ft plot at the Arthur H. Post Agronomy Farm (43°38'19.39"N, 116°14'28.86"W), an extension research station of Montana State University in Bozeman, MT. The cage structure was built using 1-inch PVC piping with the netting made using 530\(\mu\) Amber Lumite Screen (BioQuip® Products, LLC). Twelve cages were built to dimensions of 6ft x 3ft x 3ft (L x W x H) with cage locations selected randomly based on the space available within the plot and arranged in sets of three. The cages were removed from the plot after 3 weeks of sawfly release.

### Insects

Wheat stem stubble was collected in Three Forks, MT, USA (43°38'19.39"N, 116°14'28.86"W) from fields that experienced high levels of C. cinctus infestation and cutting in the prior year. Cut stubble, which contained C. cinctus larvae in diapause, were kept refrigerated between -2°C and 3°C for >100 days as required to complete obligatory larval diapause. As needed, stubs were removed from refrigeration and kept at 22-27°C for 4-5 weeks inside of 100 oz GladWare® storage containers (Glad®, Oakland, California USA). Stubs were added to cages once B. inermis stems reached six inches tall, with sawfly quantity treatments as follows: high (600 stubs), low (200 stubs), and control (0 stubs).

#### Data Collection

Stems were collected from each cage site in late August. All stems from each 6 x 3 ft cage section were collected. Each stem was sliced open using X-Acto® knives to collect data on infestation, dead larvae, live larvae, and parasitism at each internode.

## Analysis

### Overall WSS Infestation and Cutting

```{r (pf.RAW) post farm n.inf data, echo = F, message = F}
pf.raw.2022 <- read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2022/Post\ Farm\ 2022.xlsx", sheet = "n.inf")
pf.raw.2023 <- read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2023/Post\ Farm\ 2023.xlsx", sheet = "n.inf")
```
```{r (pf.inf.23) post farm inf data, echo = F}
pf.inf.23 <- read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Post\ Farm/2023/Post\ Farm\ 2023.xlsx", sheet = "inf") %>%
  rename(n.1 = 8, n.2 = 9, n.3 = 10, n.4 = 11,
         n.5 = 12, n.6 = 13, n.7 = 14, n.8 = 15,
         n.9 = 16, n.10 = 17)
```
```{r (pf.comb) post farm data combined, echo = F}
# 2022
pf.2022.site <-
pf.raw.2022 %>%
  group_by(treatment, site) %>%
  reframe(site, treatment,
            p.inf = sum(iv, ir, ix)/sum(uv, ur, iv, ir, ix),
            cut = sum(cut)/sum(total),
            total.stems = sum(iv, ir, ix, uv, ur)) %>%
  unique() %>%
  pivot_longer(-c(treatment, total.stems, site),
               names_to = "type",
               values_to = "percent") %>%
  mutate(year = "2022")

# 2023
pf.2023.site <-
pf.raw.2023 %>%
  group_by(treatment, site) %>%
  reframe(treatment, site, 
          t.inf = sum(n.iv, n.ir, n.ix),
          p.inf = sum(n.iv, n.ir, n.ix)/sum(n.uv, n.ur, 
                                            n.iv, n.ir, n.ix),
          cut = sum(cut)/sum(total),
          total.stems = sum(n.iv, n.ir, n.ix, n.uv, n.ur)) %>%
  distinct() %>%
  pivot_longer(-c(treatment, total.stems, site),
               names_to = "type",
               values_to = "percent") %>%
  filter(!type == "t.inf") %>%
  arrange(type) %>%
  mutate(year = "2023")

pf.comb <-
  pf.2022.site %>%
  rbind(pf.2023.site) %>%
  group_by(treatment, year, type) %>%
  mutate(count = n(),
         average = mean(percent),
         sd = sd(percent),
         se = (sd/sqrt(count))) %>%
  reframe(treatment, year, type,
          mean = mean(percent),
          sd = mean(sd),
          se = mean(se)) %>%
  unique()

```
```{r (pf.total) pf total stems, echo = F}
# 2022 totals
pf.2022.total <-
  pf.raw.2022 %>%
  group_by(treatment) %>%
  reframe(total = sum(uv, ur, iv, ir, ix)) %>%
  mutate(year = "2022")

# 2023 totals
pf.2023.total <-
  pf.raw.2023 %>%
  group_by(treatment) %>%
  reframe(total = sum(n.uv, n.ur, n.iv, n.ir, n.ix)) %>%
  mutate(year = "2023")

pf.total.comb <-
  pf.2022.total %>%
  rbind(pf.2023.total)

```

#### Basic Table and Plot of Infestation

```{r post farm table, echo = F, results = 'asis'}
pf.23.table <-
pf.comb %>%
  filter(year == "2023") %>%
  dplyr::select(treatment, type, mean) %>%
  mutate(mean = round_based_on_size(mean)) %>%
  pivot_wider(values_from = mean,
              names_from = type,
              values_fn = list(mean = mean)) %>%
  left_join(pf.2023.total, by = "treatment")

pf.22.table <-
pf.comb %>%
  filter(year == "2022") %>%
  dplyr::select(treatment, type, mean) %>%
  mutate(mean = round_based_on_size(mean)) %>%
  pivot_wider(values_from = mean,
              names_from = type,
              values_fn = list(mean = mean))  %>%
  left_join(pf.2022.total, by = "treatment")

pf.23.table.final <-
kable(pf.23.table[c("treatment", "total", "p.inf", "cut")], 
      caption = "2023 Post Farm Infestation", 
      format = "html",
      col.names = c("Treatment", "Total Stems", "% Inf", "% Cut"),
      align = "lccc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f2")

pf.22.table.final <-
kable(pf.22.table[c("treatment", "total", "p.inf", "cut")], 
      caption = "2022 Post Farm Infestation", 
      format = "html",
      col.names = c("Treatment", "Total Stems", "% Inf", "% Cut"),
      align = "lcc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f2")

cat('<div style="display: flex;">')
cat('<div style="flex: 1;">')
cat(pf.23.table.final)
cat('</div>')
cat('<div style="flex: 1;">')
cat(pf.22.table.final)
cat('</div>')
cat('</div>')
```

```{r post farm plot, echo = F, fig.height = 2.5, fig.align = 'center'}
pf.comb <-
pf.comb %>%
  left_join(pf.total.comb, by = c("treatment", "year"))

ggplot() +
  facet_wrap(~treatment) +
  geom_bar(stat = "identity",
           position = position_dodge(),
           data = pf.comb %>%
             filter(type == "p.inf"),
           aes(year, mean, fill = "p.inf")) +
  geom_bar(stat = "identity",
           position = position_dodge(),
           data = pf.comb %>%
             filter(type == "cut"),
           aes(year, mean, fill = "cut")) +
  geom_errorbar(data = pf.comb,
                aes(x = year,
                    ymin = mean - se,
                    ymax = mean + se),
                position = position_dodge(0.9),
                width = 0.2,
                alpha = 0.5) +
  theme_bw() +
  labs(x = element_blank(),
       y = "Percentage of stems (%)",
       fill = element_blank()) +
  scale_fill_brewer(palette = "Greens",
                    labels = c("Cut", "Infested"))

```

Let's take a closer look at the number of nodes borrowed through by the larvae. Using this information we can make an arguement about the size and length of time each larvae was alive prior to dying. This can be informative in understanding how many larvae would have been left later in the season for potential parasitism.

#### Infestation Percentage/Node

```{r v.r infestation node, echo = F, warning = F, message = F, fig.height = 3.5, fig.align = 'center'}
v.r_inf <- 
pf.inf.23 %>%
  dplyr::select(site, rep, treatment, 8:19) %>%
  mutate(stem.id = row_number()) %>%
  pivot_longer(-c(stem.id, site, rep, treatment, v.r, lar),
               names_to = "node",
               values_to = "node_status") %>%
  na.omit() %>%
  group_by(stem.id, treatment, v.r, lar) %>%
  reframe(nodes.inf = n()) %>%
  group_by(treatment, v.r, lar, nodes.inf) %>%
  reframe(count = n()) %>%
  group_by(treatment, v.r) %>%
  mutate(count2 = sum(count),
         per = count/count2) %>%
  reframe(treatment, v.r, nodes.inf, per) %>%
  mutate_at(vars(nodes.inf), as.factor)

# plot
v.r_inf %>%
  ggplot() +
  geom_bar(stat = "identity",
         position = position_dodge(),
         aes(y = per, x = nodes.inf, fill = treatment)) +
  geom_line(stat = "smooth",
            aes(y = per, 
                x = nodes.inf, 
                group = treatment, 
                color = treatment),
            se = FALSE,
            alpha = 0.6,
            linewidth = 1.3,
            linetype = "solid",
            method = "loess") +
  theme_bw() +
  scale_fill_brewer(palette = "Greens") +
  scale_color_brewer(palette = "Reds") +
  facet_wrap(~v.r)

# anova
anova1 <- 
  summary(aov(per ~ treatment + nodes.inf + v.r, data = v.r_inf))

aov.df <- as.data.frame(anova1[[1]])
aov.df$`Pr(>F)` <- sprintf("%.4f", aov.df$`Pr(>F)`)
aov.df$Source <- rownames(aov.df)
rownames(aov.df) <- NULL

aov.df[,2:4] <- round(aov.df[,2:4], 3)

# anova table
kbl(aov.df %>%
      dplyr::select(Source, everything()), 
    align = "lccccc",
    format = "html",
    caption = "ANOVA Table",
    col.names = c("", "Df", "Sum Sq", "Mean Sq", "F-Value", "Pr(>F)")) %>%
  kable_styling(bootstrap_options = c("striped", "bordered", 
                                      "hover", "repsonsive"),
                full_width = FALSE) %>%
  row_spec(0, background = "#f2f2f2")

```

Based on the results of our ANOVA, we have weak evidence (Pval = `r aov.df[3,5]`) to suggest there is a difference in the number of nodes infested when considering treatment group and vegetative and reproductive stems.

```{r cumsum of infestation per node, echo = F, fig.align = "center"}
# I want to see the percentage of stems with 6+ nodes infested, then 5:6+, then 4:6+, etc.
cumsum <-
pf.inf.23 %>%
  dplyr::select(site, rep, treatment, 8:19) %>%
  mutate(stem.id = row_number()) %>%
  pivot_longer(-c(stem.id, site, rep, treatment, v.r, lar),
               names_to = "node",
               values_to = "node_status") %>%
  na.omit() %>%
  group_by(stem.id, treatment, v.r, lar) %>%
  reframe(nodes.inf = n()) %>%
  group_by(treatment, nodes.inf) %>%
  reframe(count = n()) %>%
  group_by(treatment) %>%
  mutate(count2 = sum(count),
         per = count/count2) %>%
  reframe(treatment, nodes.inf, per) %>%
  arrange(treatment, desc(nodes.inf)) %>%
  group_by(treatment) %>%
  mutate(cumsum = cumsum(per)) %>%
  mutate_at(vars(nodes.inf), as.factor) %>%
  dplyr::select(treatment, nodes.inf, cumsum) %>%
  mutate(nodes.inf = fct_rev(nodes.inf))

cumsum %>%
  mutate(cumsum = round(cumsum, 2)) %>%
  rename("Treatment" = treatment) %>%
  pivot_wider(values_from = "cumsum",
              names_from = "nodes.inf") %>%
  kable("html",
        caption = "Cumulative sum of nodes infested",
        align = "lcccccc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f2") %>%
  add_header_above(c(" ", "# of Nodes Infested" = 6))

ggplot(data = cumsum,
         aes(x = nodes.inf,
             y = cumsum,
             group = treatment, 
             color = treatment)) +
  geom_line() +
  theme_bw() +
  geom_bar(stat = "identity",
           aes(x = nodes.inf, 
               y = cumsum,
               fill = treatment),
           position = position_dodge(),
           alpha = 0.25) +
  theme_bw() +
  labs(x = "Number of Nodes Infested",
       y = "Cumulative sum % of stems",
       fill = "WSS Treatment \nLevel",
       color = "WSS Treatment \nLevel",
       title = "Smooth brome node infestation") +
  scale_x_discrete(limits = rev(levels(cumsum$cumsum)))

```

I like this plot because it shows the cumulative percentage sum of nodes infested at each node. For example, at the 3rd node in 2023, we see that close to 70% of the stems in the 'high' treatment group exhibited between 3 and 6 nodes infested. As a general thought exercise, the more nodes a larvae burrowed through, the larger the larvae. This means that if we classify a "large" larvae as one who burrowed through at least 4 nodes, our infestation data shows that 36% of the larvae in the 'high' treatment groups died as "large" larvae.

```{r larvae breakdown, echo = F}
# how many stems had live larvae in them?
# percent of stems with each larval type

larvae.type <-
pf.inf.23 %>%
  dplyr::select(site, rep, treatment, 8:19) %>%
  mutate(stem.id = row_number()) %>%
  pivot_longer(-c(stem.id, site, rep, treatment, v.r, lar),
               names_to = "node",
               values_to = "node_status") %>%
  na.omit() %>%
  group_by(treatment, lar) %>%
  reframe(count = n()) %>%
  mutate(year = "2023") %>%
  left_join(pf.total.comb, by = c("treatment", "year")) %>%
  group_by(treatment) %>%
  mutate(per = count/total) %>%
  filter(lar == "a") %>%
  mutate(per = round(per, 3))
# stems collected in late October. This is much later than if parasitoids were flying.

kable(larvae.type[c("treatment", "count", "total", "per")],
      format = "html",
      align = "c",
      col.names = c("Treatment", "Count", "Total Stems", "%"),
      caption = "Stems with Living Larvae (2023)") %>%
  kable_styling(bootstrap_options = c("striped", "bordered", 
                                      "hover", "repsonsive"),
                full_width = FALSE) %>%
  row_spec(0, background = "#f2f2f2")
```

Interesting to see that `r larvae.type[1,6]*100`% of stems in the 'high' treatment groups and `r larvae.type[2,6]*100`% of stems in the 'low' treatment group had living WSS larvae. The stems for this experiement were harvested in late October, much later in the season when we would expect the second generation parasitoids to be searching for a host.

<hr>

# MONTANA FIELD SITES

## Methods

To assess C cinctus infestation, larval mortality, and parasitoid prevalence within B. inermis and adjacent wheat fields, sites were chosen that had the best chance of accurately representing conditions throughout the Golden Triangle, a region of north-central Montana known for its dryland grain production. Samples were collected in both early July and late August in 2021, 2022, and 2023 from sites in Big Sandy, Moccasin, and Amsterdam, MT, USA. Sample locations were chosen by identifying fields with consistent wheat production that were adjacent to healthy B. inermis stands, either situated between fields or between fields and roadways. Polygons containing B. inermis were set up along the periphery of adjacent wheat fields, each covering an area of approximately 100 meters. Four collection squares of 1ft x 1ft were randomly selected within each polygon during both collection events each year. All stems within each 1 x 1 ft square were collected. The stems were then processed and split in the laboratory to assess infestation, live eggs, dead eggs, dead larvae, live larvae, parasitism, and cutting.

## Analysis

```{r (raw) brome data, echo = F}
raw.2022 <- 
  read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Field Sites/2022/2022 FS Stem Data (2).xlsx", sheet = "n.infested") %>%
  mutate(year = "2022") %>%
  rename("site" = site.cap)

raw.2023 <-
read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Field Sites/2023/2023\ Field Site\ Data.xlsx", sheet = "n.infested") %>%
  mutate(year = "2023") 
```
```{r (stubble.####) stubble data, echo = F}
stubble.2023 <-
  read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Field Sites/2023/2023\ Field Site\ Data.xlsx", 
           sheet = "stubble") %>%
  dplyr::select(-name) %>%
  na.omit() %>%
  mutate(year = "2023")

stubble.2022 <-
  read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/BWW/Field/Field Sites/2022/2022 FS Stem Data (2).xlsx", 
             sheet = "stubble") %>%
  mutate(year = "2022")

```
```{r (joined.stubble) stubble join, echo = F}
# the below code joins the 2022 and 2023 stubble data frames
joined.stubble <-
stubble.2023 %>%
  dplyr::select(site, location, distance,
                inf.nc, cut, cocoon, eh, total, year) %>%
  mutate(collection = "fall") %>% 
  rbind(stubble.2022 %>%
          dplyr::select(site, year, location, distance, collection,
                        inf.nc, cut, cocoon, eh, total)) %>%
  # remove N sites and re add as NS and NN for example
  group_by(site, location, year, distance, collection) %>%
  mutate(inf = sum(cut, inf.nc, cocoon, eh)) %>%
  # first, we remove the site arguement by grouping by everything else
  group_by(site, location, year, distance, collection) %>%
  dplyr::select(-c(inf.nc)) %>%
  rename("e" = eh,
         "c" = cocoon) %>%
  mutate(p.inf = inf/total,
         p.par = (c + e)/inf,
         p.cut = cut/inf,
         poten = (cut + e + c)/inf)

# duplicate the wheat data to match up with brome data locations
# we are duplicating some of the stubble data so that we have stubble data
# for the brome sections that are across the road from stubble data
joined.stubble <-
  joined.stubble %>%
  bind_rows(joined.stubble %>%
              filter(location == "n") %>%
              mutate(location = "ns")) %>%
  bind_rows(joined.stubble %>%
              filter(location == "n") %>%
              mutate(location = "nn")) %>%
  filter(!location == "n") %>%
  bind_rows(joined.stubble %>%
              filter(location == "har") %>%
              mutate(location = "sun"))

```
```{r (n.inf) brome join, echo = F}
n.inf <-
raw.2023 %>%
  dplyr::select(site, year, location, collection, 
              n.uv, n.ur, n.iv, n.ir, n.ix, 
              cut, e, c)  %>%
  bind_rows(raw.2022 %>%
              rename(n.uv = uv,
                     n.ur = ur,
                     n.iv = iv,
                     n.ir = ir,
                     n.ix = ix) %>%
              dplyr::select(site, year, location, collection, 
                    n.uv, n.ur, n.iv, n.ir, n.ix, 
                    cut, e, c)) %>%
  group_by(site, location, collection, year) %>%
  mutate(inf = sum(n.iv, n.ir, n.ix),
         total = sum(n.uv, n.ur, n.iv, n.ir, n.ix)) %>%
  ungroup() %>%
  dplyr::select(-c(n.uv, n.ur, n.iv, n.ir, n.ix)) %>%
  mutate(p.inf = inf/total,
         p.par = (c + e)/inf,
         p.cut = cut/inf,
         poten = (cut + c + e)/inf) %>%
  pivot_longer(-c(site, location, collection, year),
               names_to = "type",
               values_to = "value") %>%
  group_by(location, collection, year, type) %>%
  distinct() %>%
  mutate(distance = "brome",
         location = case_when(
           location == "kw" ~ "k",
           location == "ke" ~ "k", 
           TRUE ~ location))
```
```{r (stubble.brome) join stubble/brome, echo = F}
# join the stubble data and the brome data
stubble.brome <-
joined.stubble %>%
  pivot_longer(-c(site, location, distance, year, collection),
               names_to = "type",
               values_to = "value") %>%
  distinct() %>%
  rbind(n.inf) %>%  # rbind the brome data
  filter(!type == "total") %>%  # remove type = "total"
  mutate(type = case_when(
    type == "c" ~ "p",
    type == "e" ~ "p",
    TRUE ~ type)) %>% # replace the e and c variables with p, for simplicity
  ungroup() %>%
  dplyr::select(-site) %>%
  group_by(location, distance, year, collection, type) %>%
  reframe(mean = mean(value),
         count = n(),
         sd = sd(value),
         se = sd/sqrt(count)) %>%
  distinct() %>%
  filter(type %in% c("p.inf", "p.cut", "p.par", "poten")) %>%
  distinct()
```

```{r, echo = F, eval = F}
stubble.brome %>%
  filter(location == "g",
         distance %in% c("20", "5"),
         year == "2023",
         collection == "fall")
```

### Big Sandy, MT 2023

#### Lochy Property - North Road

![North Road, Big Sandy, MT (2023)](https://github.com/jstrand894/Website/blob/main/nnns-min.png?raw=true)

**Note:**    
- In 2023, the field south of North Road was fallow, so no stubble infestation data was collected from the field south of North Road.

```{r NN & NS data table, echo = F}
nn.table <-
stubble.brome %>%
  filter(location %in% c("nn"),
         year == "2023") %>%
  dplyr::select(location, distance, collection, year, type, mean) %>%
  mutate(mean = round_based_on_size(mean)) %>%
  mutate_at(vars(location, distance, collection, year, type), as.factor) %>%
  pivot_wider(names_from = type,
              values_from = mean,
              values_fn = list(mean = mean)) %>%
  arrange(distance) %>%
  ungroup() %>%
  dplyr::select(-year)

kable(nn.table[, column.order2], 
      caption = "North Road (site nn) Infestation Table", 
      format = "html",
      col.names = c("Location", "Distance", "Collection", 
                    "% Inf", "Cut", "Parasitized", "Potential Cut"),
      align = "lllcccc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f2")

ns.table <-
stubble.brome %>%
  filter(location %in% c("ns"),
         year == "2023",
         distance == "brome") %>%
  dplyr::select(location, distance, collection, year, type, mean) %>%
  mutate(mean = round_based_on_size(mean)) %>%
  mutate_at(vars(location, distance, collection, year, type), as.factor) %>%
  pivot_wider(names_from = type,
              values_from = mean,
              values_fn = list(mean = mean)) %>%
  arrange(distance) %>%
  ungroup() %>%
  dplyr::select(-year)

kable(ns.table[, column.order2], 
      caption = "North Road (site ns) Infestation Table", 
      format = "html",
      col.names = c("Location", "Distance", "Collection", 
                    "% Inf", "Cut", "Parasitized", "Potential Cut"),
      align = "lllcccc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f2") 
```

```{r NN and NS plot, echo = F, fig.height = 3.5, fig.width = 10, warning = F}
stubble.brome %>%
  filter(location %in% c("nn", "ns"),
         year == "2023",
         collection == "fall") %>%
  ggplot(aes(x = type, y = mean, fill = distance)) +
  geom_bar(stat = "identity",
           position = position_dodge()) +
  geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                position = position_dodge(0.9),
                width = 0.2,
                alpha = 0.5) +
  theme_bw() +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Purples",
                    labels = c("Adjacet wheat: 20 m", "Adjacet wheat: 5 m", "Smooth brome")) +
  labs(y = "Percent (%)",
       x = element_blank(),
       title = "North Road Infestation") +
  scale_x_discrete(labels = c("p.cut" = "Cut",
                              "p.inf" = "Infested",
                              "p.par" = "Parasitized",
                              "poten" = "Potential Cut")) +
  facet_wrap(~location)
```

We saw very low levels of infestation in 2023 within the wheat. This was likely due a cold, mild spring that led to late stem development. Because the wheat was not well developed, the sawflies did not infest. This appears to have led to high infestation within the smooth brome. Will be interesting to compare to 2022 when stem elongation followed normal expected timelines.

#### Lochy Property - Gardiner Road

![Gardiner Road, Big Sandy, MT (2023)](https://github.com/jstrand894/Website/blob/main/gardiner%20road-min.png?raw=true)

In 2022, Gardiner road underwent a major reconstruction with both ditches being regraded. This severely diminished the amount smooth brome found along the edge of the roadway.

```{r G table, echo = F}
g.table <-
stubble.brome %>%
  filter(location %in% c("g"),
         year == "2023") %>%
  dplyr::select(location, distance, collection, year, type, mean) %>%
  mutate(mean = round_based_on_size(mean)) %>%
  mutate_at(vars(location, distance, collection, year, type), as.factor) %>%
  pivot_wider(names_from = type,
              values_from = mean,
              values_fn = list(mean = mean)) %>%
  arrange(distance) %>%
  ungroup() %>%
  dplyr::select(-c(year, location))

kable(g.table[, column.order], caption = "Gardiner Road Infestation Table", 
      format = "html",
      col.names = c("Distance", "Collection", 
                    "% Inf", "Cut", "Parasitized", "Potential Cut"),
      align = "lllcccc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f2")
```

```{r G plot, echo = F, fig.height = 3.5, fig.align = 'center', warning = F}
stubble.brome %>%
  filter(location %in% c("g"),
         year == "2023",
         collection == "fall") %>%
  ggplot(aes(x = type, y = mean, fill = distance)) +
  geom_bar(stat = "identity",
           position = position_dodge()) +
  geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                position = position_dodge(0.9),
                width = 0.2,
                alpha = 0.5) +
  theme_bw() +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Purples",
                    labels = c("Adjacet wheat: 20 m", "Adjacet wheat: 5 m", "Smooth brome")) +
  labs(y = "Percent (%)",
       x = element_blank(),
       title = "Gardiner Rd Infestation") +
  scale_x_discrete(labels = c("p.cut" = "Cut",
                              "p.inf" = "Infested",
                              "p.par" = "Parasitized",
                              "poten" = "Potential Cut"))
```

Similar results to North Road seen here. Low infestation in the adjacent wheat field was due to the cool spring and lack of infestable stems during the sawfly flight. This shows that even in odd years like 2023, sawflies are still present in the smooth brome, meaning parasitoids will still have larval hosts.

### Moccasin, MT 2023

#### Sunset & Harvest Road Intersection

![Harvest & Sunset Road Instersection, Big Sandy, MT (2023)](https://github.com/jstrand894/Website/blob/main/harvest%20road-min.png?raw=true)

```{r SUN/HAR data tables, echo = F}
sun.table <-
stubble.brome %>%
  filter(location %in% c("sun"),
         year == "2023") %>%
  dplyr::select(location, distance, collection, year, type, mean) %>%
  mutate(mean = round_based_on_size(mean)) %>%
  mutate_at(vars(location, distance, collection, year, type), as.factor) %>%
  pivot_wider(names_from = type,
              values_from = mean,
              values_fn = list(mean = mean)) %>%
  arrange(distance) %>%
  ungroup() %>%
  dplyr::select(-year)

kable(sun.table[, column.order2], caption = "Sunset Road Infestation Table", 
      format = "html",
      col.names = c("Location", "Distance", "Collection", 
                    "% Inf", "Cut", "Parasitized", "Potential Cut"),
      align = "lllcccc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f2")

har.table <-
stubble.brome %>%
  filter(location %in% c("har"),
         year == "2023") %>%
  dplyr::select(location, distance, collection, year, type, mean) %>%
  mutate(mean = round_based_on_size(mean)) %>%
  mutate_at(vars(location, distance, collection, year, type), as.factor) %>%
  pivot_wider(names_from = type,
              values_from = mean,
              values_fn = list(mean = mean)) %>%
  arrange(distance) %>%
  ungroup() %>%
  dplyr::select(-year)

kable(har.table[, column.order2], caption = "Harvest Road Infestation Table", 
      format = "html",
      col.names = c("Location", "Distance", "Collection", 
                    "% Inf", "Cut", "Parasitized", "Potential Cut"),
      align = "lllcccc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f2")
```

```{r SUN/HAR plot, echo = F, warning = F, fig.align = 'center', fig.height = 3.5, fig.width = 10}
stubble.brome %>%
  filter(location %in% c("sun", "har"),
         year == "2023",
         collection == "fall") %>%
  ggplot(aes(x = type, y = mean, fill = distance)) +
  geom_bar(stat = "identity",
           position = position_dodge()) +
  geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                position = position_dodge(0.9),
                width = 0.2,
                alpha = 0.5) +
  theme_bw() +
  theme(legend.title = element_blank()) +
  guides(color = guide_legend(title.position = "top", 
                              label.padding = unit(10, "lines"))) +
  scale_fill_brewer(palette = "Purples",
                    labels = c("Adjacet wheat: 20 m", "Adjacet wheat: 5 m", "Smooth brome")) +
  labs(y = "Percent (%)",
       x = element_blank(),
       title = "Sunset & Harvest Rd Infestation") +
  scale_x_discrete(labels = c("p.cut" = "Cut",
                              "p.inf" = "Infested",
                              "p.par" = "Parasitized",
                              "poten" = "Potential Cut")) +
  facet_wrap(~location)
```

#### Kolin Road

![Kolin Road, Moccasin, MT (2023)](https://github.com/jstrand894/Website/blob/main/kolin%20road-min.png?raw=true)

```{r K data table, echo = F}
k.table <-
stubble.brome %>%
  filter(location %in% c("k"),
         year == "2023") %>%
  dplyr::select(location, distance, collection, year, type, mean) %>%
  mutate(mean = round_based_on_size(mean)) %>%
  mutate_at(vars(location, distance, collection, year, type), as.factor) %>%
  pivot_wider(names_from = type,
              values_from = mean,
              values_fn = list(mean = mean)) %>%
  arrange(distance) %>%
  ungroup() %>%
  dplyr::select(-year)

kable(k.table[, column.order], caption = "Kolin Road Infestation Table", 
      format = "html",
      col.names = c("Distance", "Collection", 
                    "% Inf", "Cut", "Parasitized", "Potential Cut"),
      align = "lllcccc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f2")
```

```{r K plot, echo = F, warning = F, fig.height = 3, fig.align = 'center'}
stubble.brome %>%
  filter(location %in% c("k"),
         year == "2023",
         collection == "fall") %>%
  ggplot(aes(x = type, y = mean, fill = distance)) +
  geom_bar(stat = "identity",
           position = position_dodge()) +
  geom_errorbar(aes(ymin = mean - se,
                    ymax = mean + se),
                position = position_dodge(0.9),
                width = 0.2,
                alpha = 0.5) +
  theme_bw() +
  theme(legend.title = element_blank()) +
  scale_fill_brewer(palette = "Purples",
                    labels = c("Adjacet wheat: 20 m", "Adjacet wheat: 5 m", "Smooth brome")) +
  labs(y = "Percent (%)",
       x = element_blank(),
       title = "Kolin Rd Infestation") +
  scale_x_discrete(labels = c("p.cut" = "Cut",
                              "p.inf" = "Infested",
                              "p.par" = "Parasitized",
                              "poten" = "Potential Cut")) 
```