### 2024-02-29 msu-masters.qmd

---
title: "MSU Masters"

---
<figure>
  <img src="img/msu-masters/cages.heic" alt="Montana State University" 
  width = 250" height = "200">
</figure>

<figure>
  <img src="img/msu-masters/vcs.heic" alt="Montana State University" 
  width = 200" height = "300">
</figure>

<figure>
  <img src="img/msu-masters/field.heic" alt="Montana State University" 
  width = 250" height = "200">
</figure>