### 2024-02-29 projects.qmd

---
title: "Projects"
editor_options: 
  chunk_output_type: console
---
::: {style="display: flex;"}

<figure style="margin-right: 20px; display: flex; flex-direction: column; align-items: center;">
  <a href="msu-masters.qmd">
    <img src="img/msu-logo.jpg" alt="Montana State University" width="200" height="200">
  </a>
  <figcaption style="text-align: center;">Montana State University</figcaption>
</figure>

<figure style="margin-right: 20px; display: flex; flex-direction: column; align-items: center;">
  <a href="bballref.qmd">
    <img src="img/bballref-logo.jpeg" alt="Basketball Reference Tool" width="200" height="200">
  </a>
  <figcaption style="text-align: center;">Basketball Reference Tool</figcaption>
</figure>

:::