### 2024-02-29 tmr50k.qmd

---
title: "Teton Mountain Runs 50K"
---

::: {style="display: block;"}

```{r get images, echo = F, results = 'asis'}
image_files <- list.files("img/tmr50k", pattern = ".JPG|.png|.gif|.jpg|.jpeg", full.names = TRUE)

for (file in image_files) {
  cat("![](",file, ")\n  \n")
}
```

:::