### 2024-03-01 cv.qmd

---
layout: page
title: "Cirriculum vitae"
execute:
  freeze: true  # never re-render during project render
engine: knitr 
resources:
  - "Aboutme/jackson_strand_updated_cv.pdf"
cv:
  pdf: "Aboutme/jackson_strand_updated_cv.pdf"
---

```{css echo=FALSE}
.embed-container {
    position: relative;
    padding-bottom: 129%;
    height: 0;
    overflow: hidden;
    max-width: 100%;
}
.embed-container iframe,
.embed-container object,
.embed-container embed {
    position: absolute;
    top: 0;
    left: 0;
    width: 100%;
    height: 100%;
}
```

```{=html}
<p class="text-center">
  <a class="btn btn-primary btn-lg cv-download" href="`r rmarkdown::metadata$cv$pdf`" target="_blank">
    <i class="fa-solid fa-file-arrow-down"></i>&ensp;Download current CV
  </a>
</p>
<div class="embed-container">
  <iframe src="`r rmarkdown::metadata$cv$pdf`#toolbar=0" style="border: 0.5px;width: 100%;"></iframe>
</div>
```