### 2024-03-05 aboutme.qmd
Update website 4/5/24

---
title: "Jackson Strand"
about:
  id: me
  template: trestles
  links:
    - icon: linkedin
      text: LinkedIn
      href: https://linkedin.com/in/jacksonstrand
    - icon: instagram
      text: Instagram
      href: https://www.instagram.com/jstrand894/
    - icon: github
      text: GitHub
      href: https://github.com/jstrand894 
  image: img/aboutme/deadhorse.jpeg

sidebar: false
---

:::{#me}

I am currently a Master's graduate student at Montana State University studying entomology. I live in Bozeman, MT with my wife Grace.

```{=html}
<div  style="margin: 30px; text-align: center;">
  <a class="btn btn-primary" href="cv.qmd" role="button" style="padding: 15px 30px;">View my CV</a>
</div>
```

![Me and Grace at Glacier National Park, 2023](img/aboutme/iceberg.JPG)

:::

```{=html}
<!DOCTYPE html>
<html lang="en">
<head>
  <meta charset="UTF-8">
  <meta name="viewport" content="width=device-width, initial-scale=1.0">
  <title>Image Gallery</title>
  <style>
    .gallery {
      display: grid;
      grid-template-columns: repeat(auto-fill, minmax(200px, 1fr));
      grid-gap: 10px;
    }
    .gallery img {
      width: 100%;
      height: auto;
      cursor: pointer;
      transition: transform 0.3s ease;
    }
    .gallery img:hover {
      transform: scale(1.05); /* Slightly increase size on hover */
    }
    .expanded-container {
      position: fixed;
      top: 0;
      left: 0;
      width: 100%;
      height: 100%;
      background-color: rgba(0, 0, 0, 0.8); /* Semi-transparent background */
      display: none; /* Initially hidden */
      justify-content: center;
      align-items: center;
      z-index: 9999;
    }

    .expanded {
      max-width: 80%;
      max-height: 80%;
      cursor: pointer;
    }
  </style>
</head>
<body>

<div class="gallery">
  <img src="img/bugs/blueflowers.JPG" alt="Image 1" onclick="expandImage(this)">
  <img src="img/bugs/brownmoth.JPG" alt="Image 2" onclick="expandImage(this)">
  <img src="img/bugs/butterfly-purpleflower.JPG" alt="Image 3" onclick="expandImage(this)">
  <!-- Add more images as needed -->
</div>

<div class="expanded-container" onclick="closeExpanded()">
  <img class="expanded" id="expanded-image" onclick="navigateGallery(event)">
</div>

<script>
  var currentImageIndex = 0;
  var galleryImages = document.querySelectorAll('.gallery img');

  function expandImage(img) {
    currentImageIndex = Array.from(galleryImages).indexOf(img);
    document.getElementById('expanded-image').src = img.src;
    document.querySelector('.expanded-container').style.display = 'flex';
  }

  function closeExpanded() {
    document.querySelector('.expanded-container').style.display = 'none';
  }

  function navigateGallery(event) {
    // Calculate the click position relative to the expanded image
    var clickX = event.pageX - event.target.getBoundingClientRect().left;
    var imageWidth = event.target.offsetWidth;

    // Determine if the click was on the left or right side of the image
    if (clickX < imageWidth / 2) {
      // Clicked on the left side: navigate to the previous image
      currentImageIndex = (currentImageIndex - 1 + galleryImages.length) % galleryImages.length;
    } else {
      // Clicked on the right side: navigate to the next image
      currentImageIndex = (currentImageIndex + 1) % galleryImages.length;
    }

    // Update the expanded image source
    document.getElementById('expanded-image').src = galleryImages[currentImageIndex].src;
  }
</script>

</body>
</html>

```