### 2024-03-05 index.qmd

---
about: 
  template: jolla
  id: about-block
  image: "img/aboutme/crescent.HEIC"
  links:
    - icon: linkedin
      text: LinkedIn
      href: https://linkedin.com/in/jacksonstrand
    - icon: instagram
      text: Instagram
      href: https://www.instagram.com/jstrand894/
    - icon: github
      text: GitHub
      href: https://github.com/jstrand894 
---
::: {#about-block}
:::

# Welcome to my website!

My name is Jackson Strand. I am currently a Master's graduate student at Montana State University studying entomology. I live in Bozeman, MT with my wife Grace.

```{=html}
<div  style="margin: 30px; text-align: center;">
  <a class="btn btn-primary" href="cv.qmd" role="button" style="padding: 15px 30px;">View my CV</a>
</div>
```

![Me and Grace at Glacier National Park, 2023](img/aboutme/iceberg.JPG)