### 2024-03-05 running.qmd

---
title: "Running Pictures"
---
::: {style="display: flex;"}
<figure style="margin-right: 20px; display: flex; flex-direction: column; align-items: center;">
  <a href="deadhorse.qmd">
    <img src="img/deadhorse/deadhorse-top.JPG" alt="Deadhorse Ultra" width="300" height="200">
  </a>
  <figcaption style="text-align: center;">Deadhorse Ultra, 2023</figcaption>
</figure>

<figure style="margin-right: 20px; display: flex; flex-direction: column; align-items: center;">
  <a href = "tmr50k.qmd">
    <img src="img/tmr50k/tmr50k-top.JPG" alt="Teton Mountain Runs 50K" 
  width = 300" height = "200">
  </a>
  <figcaption style="text-align:center;">Teton Mountain Runs, 2023</figcaption>
</figure>

:::