### 2024-03-05 save.qmd

---
title: "save"
---

# box with shadow when you hover

```{=html}
<!DOCTYPE html>
<html lang="en">
<head>
<meta charset="UTF-8">
<meta name="viewport" content="width=device-width, initial-scale=1.0">
<title>Box with Rounded Edges</title>
<style>
  .container {
    width: 300px;
    height: 250px; /* Set container height */
    border: 0.5px solid #ccc; /* Light grey color */
    border-radius: 8px; /* Less rounded edges */
    overflow: hidden; /* Hide overflow content */
    position: relative;
    display: inline-block;
    transition: box-shadow 0.3s ease; /* Transition for box-shadow */
    box-shadow: 0 2px 5px rgba(0, 0, 0, 0.1); /* Slight shadow by default */
    text-align: center; /* Center content horizontally */
    text-decoration: none; /* Remove underline */
  }

  .container:hover {
    box-shadow: 0 5px 10px rgba(0, 0, 0, 0.1); /* Increased shadow on hover */
  }

  .content {
    width: 100%;
    height: 100%;
    padding: 10px;
    box-sizing: border-box;
  }

  .content img {
    width: 100%;
    height: auto;
    display: block;
    margin: 0 auto; /* Center the image horizontally */
    object-fit: cover; /* Crop the image to cover container */
  }

  .text {
    margin-top: 10px;
    text-align: center; /* Center text within the box */
    text-decoration: none;
  }

  .text h1 {
    font-size: 24px; /* Larger text size */
    margin-top: 1px;
    margin-bottom: 5px; /* Adjust margin between large and small text */
    color: black;
  }
  .text h1:hover {
    color: #2068DC; /* Change text color to blue on hover */
  }
</style>
</head>
<body>
  <a href="https://www.example.com" target="_blank" class="container">
    <div class="content">
      <img src="img/msu-masters/wss.JPG" alt="Image"> <!-- Center the image within the box -->
      <div class="text">
        <h1>Wheat Stem Sawfly</h1>
        <!-- <h2></h2> -->
      </div>
    </div>
  </a>
</body>
</html>

```