### 2024-03-05 weatherdata.qmd

---
title: "Weather Data MT"
author: "Jackson Strand"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(glue)
library(lubridate)
library(ggplot2)
```

```{r}
inventory_url <- "https://www.ncei.noaa.gov/pub/data/ghcn/daily/ghcnd-inventory.txt"

inventory <- read_table(inventory_url,
           col_names = c("station", "lat", "lon", "variable", "start", "end"))

inventory

my_lat <- 45.661966103346735 * 2 * pi/360
my_long <- -111.04939878991246 * 2 * pi/360

my_station <- 
  inventory %>%
    mutate(lat_r = lat *2 *pi/360,
           lon_r = lon *2 *pi/360,
           d = 1.609344 * 3963 * acos((sin(lat_r) * sin(my_lat)) + cos(lat_r) * cos(my_lat) * cos(my_long - lon_r))) %>%
    filter(start < 1960 & end > 2020) %>%
    top_n(n = -1, d) %>%
    distinct(station) %>%
    pull(station)

station_daily <- glue("https://ncei.noaa.gov/pub/data/ghcn/daily/by_station/{my_station}.csv.gz")

local_weather <- 
read_csv(station_daily,
         col_names = c("station", "date", "variable", "value", "a", "b", "c", "d")) %>%
  dplyr::select(station, date, variable, value) %>%
  pivot_wider(names_from = "variable", values_from = "value",
                values_fill = 0) %>%
  dplyr::select(date, TMAX, PRCP, SNOW) %>%
  mutate(date = ymd(date))

```

```{r}
bozeman_weather <-
  local_weather %>%
  filter(date > "2000-01-01")

ggplot(data = bozeman_weather) +
  geom_line(aes(x = date, y = TMAX)) +
  theme_bw()
```
title: "Montana Weather Data"
---