### 2024-03-06 eag.qmd
commit 3/6/24

---
title: "eag"
editor_options: 
  chunk_output_type: console
---

```{r library, echo = F, message = F}
library(dplyr)
library(ggplot2)
library(readxl)
library(plotly)
```
```{r scale values, echo = F}
# Define the original range
min_original <- 1
max_original <- 55680

# Define the target range
min_target <- 1.823
max_target <- 10.02

# Function to scale the values
scale_values <- function(x) {
  ((x - min_original) / (max_original - min_original)) * (max_target - min_target) + min_target
}

```

```{r, echo = F}
eag <-
  read_excel("xlsx files/eag.xlsx", sheet = "test eag")

fid <-
  read_excel("xlsx files/eag.xlsx", sheet = "Peak")

# Generate a sequence of MeasRetTime values every 0.05
new_MeasRetTime <- seq(min(fid$MeasRetTime), 
                       max(fid$MeasRetTime), by = 0.05)

# Create a data frame with the new MeasRetTime values
new_data <- data.frame(MeasRetTime = new_MeasRetTime)

# Merge the new data frame with the original data, filling missing values with 0
filled_data <- merge(fid, new_data, by = "MeasRetTime", all = TRUE)
filled_data[is.na(filled_data)] <- 0

fid_plot <- 
filled_data %>%
  dplyr::select(MeasRetTime, Area, Height) %>%
  mutate(Height = ifelse(MeasRetTime < 3.2, 0, Height))
  # ggplot() +
  # geom_point(aes(x = MeasRetTime, y = Height)) +
  # geom_line(aes(x = MeasRetTime, y = Height)) +
  # theme_bw()
eag_plot <-
eag %>%
  mutate(row_num = row_number()) %>%
  rename("EAD" = 'EAD 1') 
ggplotly(
ggplot() +
  geom_line(data = fid_plot, aes(x = MeasRetTime, 
                                 y = Height), 
            color = "red") +
  geom_line(data = eag_plot, aes(x = scale_values(row_num)+1.2, 
                                 y = EAD * 100), 
            color = "blue") +

  scale_y_continuous(name = "EAD", sec.axis = sec_axis(~./100, name = "Height")) + 
  labs(title = "Comparison of EAD and Height",
       x = "Row Number (Scaled)", 
       y = "EAD",
       color = "Legend") +
  theme_bw() 
)

```