### 2024-03-06 toadflax_report.qmd
commit end of day

---
title: "Toadflax Analysis"
author: "Jackson Strand"
date: '`r format(Sys.Date())`'
editor_options: 
  chunk_output_type: console
toc: true
bibliography: citations/citations.bib
link-citations: true
cls: council-of-science-editors-author-date.csl
---

```{r libraries, echo = F, message = F}
library(dplyr)
library(tidyr)
library(readxl)
library(compositions) # 
library(ggplot2) # ggplot
library(vegan) # Adonis
library(tibble)
library(kableExtra) # Kable
library(plotly)
library(stats) # PCA
library(ggbiplot) # PCA biplot

```

```{r data load, echo = F}
data <-
read_excel("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Data/Toadflax/Toadflax 2023 Parent and Hybrid Volatiles.xlsx",
           sheet = "compounds") %>%
  rename("gc_file" = 1,
         "plant_id" = 2,
         "plant" = 3,
         "type" = 4,
         "Z-3-Hexenol" = 5,
         "a-Pinene" = 6,
         "Thujene" = 7,
         "Myrcene" = 8,
         "1-octen-3-ol" = 9,
         "6-methyl-3-heptanone" = 10,
         "B-pinene" = 11,
         "Z-3-Hexenyl Acetate" = 12,
         "Limonene" = 13,
         "Trans-B-ociemene" = 14,
         "Cis-B-ocimene" = 15,
         "Undecane" = 16,
         "Linalool" = 17,
         "Nonanal" = 18,
         "Octenyl Acetate" = 19,
         "Allo-ocimene" = 20,
         "Terpineol" = 21,
         "Methyl Salicylate" = 22,
         "Decanal" = 23,
         "P-anisaldehyde" = 24,
         "Undecanal" = 25,
         "2-Undecanal" = 26,
         "Copaene" = 27,
         "Cubebene" = 28,
         "Dodecanal" = 29,
         "Caryophyllene" = 30,
         "day" = 31) 
```
```{r CLR transformation, echo = F}
data.clr <-
data %>%
  select(-gc_file) %>%
  pivot_longer(-c(plant_id, plant, type, day),
               names_to = "compound",
               values_to = "value") %>%
  mutate(value = log(value + 1.01, 2)) %>%
  pivot_wider(values_from = "value",
              names_from = "compound") %>%
  unite(plant_id, plant_id, plant, type, day, sep = "_") %>%
  column_to_rownames("plant_id") %>%
  clr() %>% # run the clr!
  as.data.frame() %>%
  rownames_to_column("plant_id") %>%
  separate(plant_id, into = c("plant_id", "plant", "type", "day"),
           sep = "_") %>%
  pivot_longer(-c(plant_id, plant, type, day),
               names_to = "compound",
               values_to = "value") %>%
  mutate(value = value + 4) %>%
  pivot_wider(values_from = "value",
              names_from = "compound") %>%
  mutate_at(vars(plant, type, day), list(factor))
```
```{r compound order, echo = F}
compound.order <- c("Z-3-Hexenol",
"a-Pinene",
"Thujene",
"Myrcene",
"1-octen-3-ol",
"6-methyl-3-heptanone",
"B-pinene",
"Z-3-HexenylAcetate",
"Limonene",
"Trans-B-ociemene",
"Cis-B-ocimene",
"Undecane",
"Linalool",
"Nonanal",
"OctenylAcetate",
"Allo-ocimene",
"Terpineol",
"MethylSalicylate",
"Decanal",
"P-anisaldehyde",
"Undecanal",
"2-Undecanal",
"Copaene",
"Cubebene",
"Dodecanal",
"Caryophyllene")
```

I'm going to briefly analyze the data in two different formats.

- <u>Type:</u> groupings of YT, DT, Natural, and Synthetic
- <u>Plant:</u> groupings of each plant type - D6, Y6, D6Y6, Y6D6, HBR, RAD

Overall, the data is good. I think a little bit of pairing down of the number of compounds we are looking at could be helpful.

# Organized by type category
This section has data organized by the four grouping variables - YT, DT, Natural, and Synthetic.

## Simple Raw Boxplot Comparisons
```{r boxplot data category, echo = F}
boxplot.data.category <-
  data %>%
    pivot_longer(-c(gc_file, plant_id, plant, type, day),
               names_to = "compound",
               values_to = "value") %>%
    group_by(type, compound) %>%
    mutate(count = n(),
           average = mean(value),
           sd = sd(value),
           se = (sd/sqrt(count))) %>%
    reframe(type, compound,
            average = mean(average),
            sd = mean(sd),
            se = mean(se)) %>%
  unique() %>%
  mutate(compound = factor(compound, levels = compound.order)) %>%
  na.omit()
```

```{r boxplot comparison category (original data), echo = F, fig.height = 20, fig.width = 20}
boxplot.data.category %>%
  mutate(type = factor(type, levels = c("DT", "YT", "Natural", "Synthetic"))) %>%
  ggplot(aes(x = type, y = average, fill = type)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  geom_errorbar(aes(x = type,
                    ymin = average - se,
                    ymax = average + se),
                position = position_dodge(0.9),
                width = 0.2,
                alpha = 0.5) +
  labs(x = element_blank(),
       y = "VOC Concentration (ng/g/hr)") +
  theme(axis.text.x = element_text(angle = 45, 
                                    hjust = 1),
        legend.position = 'none',
        # legend.position = c(0.8,0.9),
        panel.background = element_rect(fill = "white"),
        axis.line = element_line(color = "black"),
        panel.grid.major.x = element_line(color = "#F2F2F2"),
        axis.text = element_text(size = 22),
        axis.title = element_text(size = 22),
        strip.text = element_text(size = 22)) +
  facet_wrap(~compound, 
             scales = "free_y",
             ncol = 3)

```

## Multivariate Analysis

Multivariate statistical analyses were conducted in R (R Core Team 2024) using RStudio (version 4.2.1) using PCA, Permutational Multivariate ANOVA (PERMANOVA), Permutational Analysis of Multivariate Dispersion (PERMDISP), and Random Forest modeling. The data was first transformed using a center log-ratio (CLR) transformation, a technique commonly used when analyzing multivariate compositional data [@chemoguide]. VOC compositions were compared between samples and types using a permutational multivariate analysis of variance (PERMANOVA, Canberra distance, 999 iterations) tests implemented in the vegan package (v 2.6-4) [@vegan]. This technique uses permutation testing to generate many different rearrangements of the data to assess if the observed differences are statistically significant [@anderson2017]. Data were visualized using ggplot2 [@ggplot]. The compounds used in this analysis were selected based on consistency of presence across samples , as well as overall quality of mass spectroscopy output.

### PCA

```{r PCA category, echo = F}
pca.results <-
data.clr %>%
  dplyr::select(-c(plant_id, plant, type, day)) %>%
  prcomp(scale = T)

scores <- as.data.frame(pca.results$x)
loadings <- as.data.frame(pca.results$rotation)
variance <- as.data.frame(pca.results$sdev^2)

data.plus.pca <-
  data.clr %>%
    cbind(scores)

# scree.plot <- 
#   plot(pca.results, type = "l", main = "Scree Plot")

PC1 <- pca.results$x[,1]
PC2 <- pca.results$x[,2]

loadings.plot <-
loadings %>%
  dplyr::select(PC1, PC2) %>%
  rownames_to_column("compound")

meta <- data.clr %>% dplyr::select(plant, type, day)
model.data <- data.frame(PC1, PC2, meta)
model.data$day <- factor(model.data$day, levels = c("1", "2", "3", "4",
                                                    "5", "6", "7", "8",
                                                    "9"))

```

Our first look at our PCA results is in the form of a scores plot. This shows us how our data is grouped by type.

```{r scores plot, echo = F}
ggplot() +
  geom_point(data = model.data,
             aes(PC1, PC2, color = type)) +
  # geom_text(aes(label = day))
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
  labs(x = "PC1", y = "PC2", 
       title = "Toadflax PCA Plot",
       color = element_blank()) +
  stat_ellipse(data = model.data,
               aes(PC1, PC2, color = type)) +
  labs(title = "Toadflax PCA Scores Plot") +
  theme_bw()
```

Next, we can look at our loadings and see where each compound is pushing the differences in our PCs.

```{r loadings plot, echo = F}
ggplot() +
  geom_segment(data = loadings.plot,
               aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.1, "inches")),
               alpha = 0.5,
               color = "red") +
  geom_text(data = loadings.plot,
            hjust = 1.1, vjust = -1.2,
            aes(PC1, PC2, label = compound),
            size = 4) +
  labs(x = "PC1", y = "PC2", title = "Toadflax PCA Loadings") +
  theme_bw() +
  scale_x_continuous(expand = c(0.1, 0.1)) +
  scale_y_continuous(expand = c(0.1, 0.1))
```

Finally, we can overlay the two plots to get an idea of how our data groups around our compounds. I found this plot a little bit busy, which is why I broke it down in the above two figures.

```{r sc.lo.plot, echo = F}
ggbiplot(pca.results, obs.scale = 1, var.scale = 1, ellipse = FALSE,
         group = data.clr$type, circle = FALSE) +
  theme_bw() +
  labs(title = "Toadflax PCA with loadings and scores")
```

Grouping our compounds by family (i.e. terpenes, aldehydes, etc.) could provide a little bit of more context.

### PERMANOVA

As stated above, PERMANOVA uses permutation testing to generate many different rearrangements of the data to assess if the observed differences are statistically significant [@anderson2017]. It is useful when considering non-normal multivariate data like our toadflax data set.

#### PERMANOVA - D6, Y6, Synthetic, Natural
```{r PERMANOVA all type, echo = F}
set.seed(1)

permanova_data <-
  data.clr %>%
    unite(id, plant_id, plant, type, day, sep = "_") %>%
    column_to_rownames("id") %>%
    na.omit() %>%
    mutate_all(~.^2) %>%
    mutate_all(~.*100000) %>%
    mutate_all(round, digits = 0)

meta_data <-
  data.clr %>%
  select(plant_id, plant, type, day)

# run PERMANOVA
adonis_results <-
adonis2(permanova_data ~ meta_data$type,
        # strata = meta_data$day,
        # strata = interaction(meta_data$day,
        #                      meta_data$plant),
        # method = "canberra",
        permutataions = 999,
        by = "terms")

kable(as.data.frame(adonis_results),
      format = "html",
      caption = "Toadflax Type PERMANOVA: D6, Y6, Synthetic, Natural",
      align = "lccccc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f9")

permanova.p <- adonis_results$`Pr(>F)`[1]

```

Our results show **strong** evidence (p-val = `r permanova.p`) of a difference between our four plant types (YT, DT, Synthetic, and Natural).

Let's check for a difference between our synthetic and natural plant types.

#### PERMANOVA - Synthetic & Natural
```{r PERMANOVA synthetic natural type, echo = F}
set.seed(1)

types <- c("Synthetic", "Natural")

permanova_data <-
  data.clr %>%
    filter(type %in% types) %>%
    unite(id, plant_id, plant, type, day, sep = "_") %>%
    column_to_rownames("id") %>%
    na.omit() %>%
    mutate_all(~.^2) %>%
    mutate_all(~.*100000) %>%
    mutate_all(round, digits = 0)

meta_data <-
  data.clr %>%
  filter(type %in% types) %>%
  select(plant_id, plant, type, day)

# run PERMANOVA
adonis_results <-
adonis2(permanova_data ~ meta_data$type,
        # strata = meta_data$day,
        # strata = interaction(meta_data$day,
        #                      meta_data$plant),
        # method = "canberra",
        permutataions = 999,
        by = "terms")

kable(as.data.frame(adonis_results),
      format = "html",
      caption = "Toadflax Type PERMANOVA: Synthetic & Natural",
      align = "lccccc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f9")

permanova.p <- adonis_results$`Pr(>F)`[1]
```
Our results show **strong** evidence (p-val = `r permanova.p`) of a difference between Natural and Synthetic plant types.

### PERMDISP & PCoA
Permutational Analysis of Multivariate Dispersions, or PERMDISP, is a multivariate statistic used to assess differences in dispersion (variation) among groups or treatments. This method is particularly useful when dealing with multivariate data where there are multiple variables measured on each sample or observation.

Principal Coordinates Analysis (PCoA), also known as Classical Multidimensional Scaling (MDS), is a multivariate statistical technique used for exploring and visualizing patterns of similarity or dissimilarity in data. PCoA is closely related to Principal Component Analysis (PCA), but whereas PCA operates on the covariance matrix of the data, PCoA works directly on a distance or dissimilarity matrix.

I ran a PERMDISP on our entire data set.

```{r PERMDISP, echo = F, warning = F, message = F}

dist <- dist(data.clr)

permdisp.results <-
betadisper(dist, data.clr$type)

permdisp.results
p_value <- anova(permdisp.results)$'Pr(>F)'[1]
if (p_value < 0.05) {
  cat("Significant differences in multivariate dispersion among groups.\n")
} else {
  cat("We found no significant difference (pval =",anova(permdisp.results)$'Pr(>F)'[1], ") in multivariate dispersion among groups.\n")
}

pcoa <-
capscale(dist ~ 1, data = data.clr)

pcoa_df <- data.frame(scores(pcoa)$sites, group = data.clr$type)

ggplot(pcoa_df, aes(x = MDS1, y = MDS2, color = group)) +
  geom_point(alpha = 0.7) +
  labs(x = "PCo1", y = "PCo2", color = "Type",
       title = "PCoA Plot") +
  theme_bw() +
  stat_ellipse()

```

Basically what all of the PERMDISP and PCoA results mean is that we are not seeing any significant different (`r anova(permdisp.results)$'Pr(>F)'[1]`) in the dispersion within each of our groups when we compare them to one another. We are seeing a very large p-value from our PERMDISP results and a considerable amount of overlap when looking at the PCoA plot.

This means that we don't need to consider differences in broad dispersional differences between our plant types.

# Organized by plant type
This section has data organized by the six plant types - D6, Y6, D6Y6, Y6D6, HBR, and RAD.
```{r boxplot data plant, echo = F}
boxplot.data.plant <-
  data %>%
    pivot_longer(-c(gc_file, plant_id, plant, type, day),
               names_to = "compound",
               values_to = "value") %>%
    group_by(plant, compound) %>%
    mutate(count = n(),
           average = mean(value),
           sd = sd(value),
           se = (sd/sqrt(count))) %>%
    reframe(plant, compound,
            mean = mean(average),
            sd = mean(sd),
            se = mean(se)) %>%
  unique() %>%
  left_join(data %>%
              dplyr::select(plant, type),
            by = "plant",
            relationship = "many-to-many") %>%
  unique() %>%
  mutate(plant = factor(plant, 
                        levels = c("D6", "Y6", 
                                   "D6Y6", "Y6D6",
                                   "HBR", "RAD"))) %>%
  mutate(compound = factor(compound, levels = compound.order)) %>%
  na.omit()
```

```{r boxplot comparison plant (original data), echo = F, fig.height = 20, fig.width = 20}
boxplot.data.plant  %>%
  ggplot(aes(x = plant, y = mean, fill = type)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  geom_errorbar(aes(x = plant,
                    ymin = mean - se,
                    ymax = mean + se),
                position = position_dodge(0.9),
                width = 0.2,
                alpha = 0.5) +
  labs(x = element_blank(),
       y = "VOC Concentration (ng/g/hr)") +
  theme(axis.text.x = element_text(angle = 45, 
                                    hjust = 1),
        legend.position = 'bottom',
        # legend.position = c(0.8,0.9),
        panel.background = element_rect(fill = "white"),
        axis.line = element_line(color = "black"),
        panel.grid.major.x = element_line(color = "#F2F2F2"),
        axis.text = element_text(size = 22),
        axis.title = element_text(size = 22),
        strip.text = element_text(size = 22),
        legend.text = element_text(size = 22),
        legend.title = element_text(size = 22)) +
  facet_wrap(~compound, 
             scales = "free_y",
             ncol = 3)

```

## Multivariate Analysis

### PCA

```{r PCA plant, echo = F}
pca.results <-
data.clr %>%
  dplyr::select(-c(plant_id, plant, type, day)) %>%
  prcomp(scale = T)

scores <- as.data.frame(pca.results$x)
loadings <- as.data.frame(pca.results$rotation)
variance <- as.data.frame(pca.results$sdev^2)

data.plus.pca <-
  data.clr %>%
    cbind(scores)

# scree.plot <- 
#   plot(pca.results, type = "l", main = "Scree Plot")

PC1 <- pca.results$x[,1]
PC2 <- pca.results$x[,2]

loadings.plot <-
loadings %>%
  dplyr::select(PC1, PC2) %>%
  rownames_to_column("compound")

meta <- data.clr %>% dplyr::select(plant, type, day)
model.data <- data.frame(PC1, PC2, meta)

```

```{r scores plot plant, echo = F}
ggplot() +
  geom_point(data = model.data,
             aes(PC1, PC2, color = plant)) +
  # geom_text(aes(label = day))
  geom_hline(yintercept = 0, color = "gray", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "gray", linetype = "dashed") +
  labs(x = "PC1", y = "PC2", 
       title = "Toadflax PCA Plot",
       color = element_blank()) +
  stat_ellipse(data = model.data,
               aes(PC1, PC2, color = plant)) +
  labs(title = bquote("Toadflax PCA Scores Plot by "*bold("Plant"))) +
  theme_bw() 
```

### PERMANOVA

Let's say, for example, that we want to see if there is a difference in the VOC profile between the D6Y6 and Y6D6 plants. We can run a PERMANOVA to assess for a significant difference.

```{r PERMANOVA D6Y6 & Y6D6, echo = F}
set.seed(1)

plants <- c("Y6D6", "D6Y6")

permanova_data <-
  data.clr %>%
    filter(plant %in% plants) %>%
    unite(id, plant_id, plant, type, day, sep = "_") %>%
    column_to_rownames("id") %>%
    na.omit() %>%
    mutate_all(~.^2) %>%
    mutate_all(~.*100000) %>%
    mutate_all(round, digits = 0)

meta_data <-
  data.clr %>%
  filter(plant %in% plants) %>%
  select(plant_id, plant, type, day)

# run PERMANOVA
adonis_results <-
adonis2(permanova_data ~ meta_data$plant,
        # strata = meta_data$plant,
        # strata = interaction(meta_data$day,
        #                      meta_data$plant),
        method = "canberra",
        permutataions = 999,
        by = "terms")

kable(as.data.frame(adonis_results),
      format = "html",
      caption = "Toadflax Type PERMANOVA",
      align = "lccccc") %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "bordered", "hover")) %>%
  row_spec(0, background = "#f2f2f9")

permanova.p <- adonis_results$`Pr(>F)`[1]
```

Our results show **no** evidence (p-val = `r permanova.p`) of a difference between volatile profiles between Y6D6 and D6Y6 plants.

<!-- Let's look to see if there are differences between the types individually.  -->

<!-- ```{r basic ANOVA, echo = F} -->
<!-- anova.data <- -->
<!--   data.clr %>% -->
<!--   dplyr::select(-c(plant_id, plant)) -->

<!-- compounds <- -->
<!-- anova.data[,3:26] %>% -->
<!--   colnames() -->

<!-- manova(cbind(compounds) ~ type + day, data = anova.data$type) -->
<!-- ``` -->