### 2024-03-09 msu-masters.qmd
updated website

---
title: "MSU Masters"

---
```{=html}
<!DOCTYPE html>
<html lang="en">
<head>
<meta charset="UTF-8">
<meta name="viewport" content="width=device-width, initial-scale=1.0">
<title>Box with Rounded Edges</title>
<style>
  body {
    margin: 0; /* Remove default margin */
    padding: 0; /* Remove default padding */
  }

  .container {
    width: calc(100% - 20px); /* Adjusted width to fit the screen and account for margins */
    max-width: 600px; /* Limiting maximum width */
    height: auto; /* Set container height */
    border: 0.5px solid #ccc; /* Light grey color */
    border-radius: 8px; /* Less rounded edges */
    overflow: hidden; /* Hide overflow content */
    margin-bottom: 20px; /* Add margin to create space between rows */
    margin-right: auto; /* Push containers to the left */
    margin-left: 10px; /* Adjust margin between the left side of the screen */
    display: flex; /* Use flexbox for layout */
    align-items: center; /* Center items vertically */
    text-decoration: none; /* Remove underline */
  }

  .container:last-child {
    margin-bottom: 0; /* Remove margin from the last box */
  }

  .container:hover {
    box-shadow: 0 10px 20px rgba(0, 0, 0, 0.1); /* Increased shadow on hover */
  }

  .image-container {
    padding: 20px;
    box-sizing: border-box;
  }

  .image-container img {
    width: 80px; /* Adjusted width for smaller image */
    height: auto;
    display: block;
    border-radius: 8px; /* Rounded corners for image */
  }

  .content {
    padding: 20px;
    box-sizing: border-box;
    flex: 1; /* Take remaining space */
  }

  .text h1 {
    font-size: 18px; /* Adjusted text size */
    margin-top: 0;
    margin-bottom: 5px; /* Adjust margin between large and small text */
    color: black;
  }
  .text h1:hover {
    color: #2068DC; /* Change text color to blue on hover */
  }

  .text p {
    font-size: 14px;
    margin-top: 5px;
    color: #666;
    text-align: justify; /* Align text */
  }

  @media (max-width: 768px) {
    .container {
      width: calc(100% - 20px); /* Adjusted width to fit the screen */
    }
  }
</style>
</head>
<body>
  <a href="wss.qmd" class="container">
    <div class="image-container">
      <img src="img/msu-masters/oviposition.JPG" alt="Image">
    </div>
    <div class="content">
      <div class="text">
        <h1>Wheat Stem Sawfly</h1>
        <p>Understanding the importance of smooth brome in promoting a Braconid parasitoid that targets the Wheat Stem Sawfly.</p>
      </div>
    </div>
  </a>
  <a href="locoweed-report.qmd" class="container">
    <div class="image-container">
      <img src="img/msu-masters/loco-bee2.jpg" alt="Image">
    </div>
    <div class="content">
      <div class="text">
        <h1>Locoweed Volatiles</h1>
        <p>Comparison of volatile organic compounds profiles of <em>Oxytropis</em> plants that contained the commensal endophyte <em>Alternaria oxytropis</em> and those that did not.</p>
      </div>
    </div>
  </a>
  <a href="toadflax_report.qmd" class="container">
    <div class="image-container">
      <img src="img/msu-masters/toadflax.jpeg" alt="Image">
    </div>
    <div class="content">
      <div class="text">
        <h1>Toadflax</h1>
        <p></p>
      </div>
    </div>
  </a>
</body>
</html>

```