### 2024-03-09 save.qmd

---
title: "save"
editor_options: 
  chunk_output_type: console
---

# box with shadow when you hover

```{=html}
<!DOCTYPE html>
<html lang="en">
<head>
<meta charset="UTF-8">
<meta name="viewport" content="width=device-width, initial-scale=1.0">
<title>Box with Rounded Edges</title>
<style>
  .container {
    width: 300px;
    height: 250px; /* Set container height */
    border: 0.5px solid #ccc; /* Light grey color */
    border-radius: 8px; /* Less rounded edges */
    overflow: hidden; /* Hide overflow content */
    position: relative;
    display: inline-block;
    transition: box-shadow 0.3s ease; /* Transition for box-shadow */
    box-shadow: 0 2px 5px rgba(0, 0, 0, 0.1); /* Slight shadow by default */
    text-align: center; /* Center content horizontally */
    text-decoration: none; /* Remove underline */
  }

  .container:hover {
    box-shadow: 0 5px 10px rgba(0, 0, 0, 0.1); /* Increased shadow on hover */
  }

  .content {
    width: 100%;
    height: 100%;
    padding: 10px;
    box-sizing: border-box;
  }

  .content img {
    width: 100%;
    height: auto;
    display: block;
    margin: 0 auto; /* Center the image horizontally */
    object-fit: cover; /* Crop the image to cover container */
  }

  .text {
    margin-top: 10px;
    text-align: center; /* Center text within the box */
    text-decoration: none;
  }

  .text h1 {
    font-size: 24px; /* Larger text size */
    margin-top: 1px;
    margin-bottom: 5px; /* Adjust margin between large and small text */
    color: black;
  }
  .text h1:hover {
    color: #2068DC; /* Change text color to blue on hover */
  }
</style>
</head>
<body>
  <a href="https://www.example.com" target="_blank" class="container">
    <div class="content">
      <img src="img/msu-masters/wss.JPG" alt="Image"> <!-- Center the image within the box -->
      <div class="text">
        <h1>Wheat Stem Sawfly</h1>
        <!-- <h2></h2> -->
      </div>
    </div>
  </a>
</body>
</html>

```

# how to see what the default colors are in R
```{r}
library(scales)

for(i in 1:6){
  print(hue_pal()(i))
}
hue_pal()(4)
show_col(hue_pal()(6))

```

# menu grid of boxes for selection

```{=html}
<!-- <!DOCTYPE html> -->
<!-- <html lang="en"> -->
<!-- <head> -->
<!-- <meta charset="UTF-8"> -->
<!-- <meta name="viewport" content="width=device-width, initial-scale=1.0"> -->
<!-- <title>Box with Rounded Edges</title> -->
<!-- <style> -->
<!--   .container { -->
<!--     width: 1000px; -->
<!--     height: auto; /* Set container height */ -->
<!--     border: 0.5px solid #ccc; /* Light grey color */ -->
<!--     border-radius: 8px; /* Less rounded edges */ -->
<!--     overflow: hidden; /* Hide overflow content */ -->
<!--     position: relative; -->
<!--     display: inline-block; -->
<!--     transition: box-shadow 0.15s ease; /* Transition for box-shadow */ -->
<!--     box-shadow: 0 2px 5px rgba(0, 0, 0, 0.1); /* Slight shadow by default */ -->
<!--     text-align: center; /* Center content horizontally */ -->
<!--     text-decoration: none; /* Remove underline */ -->
<!--     margin-right: 20px; /* Add margin to create space between boxes */ -->
<!--   } -->

<!--   .container:last-child { -->
<!--     margin-right: 0; /* Remove margin from the last box */ -->
<!--   } -->

<!--   .container:hover { -->
<!--     box-shadow: 0 10px 20px rgba(0, 0, 0, 0.1); /* Increased shadow on hover */ -->
<!--   } -->

<!--   .content { -->
<!--     width: 100%; -->
<!--     height: 100%; -->
<!--     padding: 20px; -->
<!--     box-sizing: border-box; -->
<!--   } -->

<!--   .content img { -->
<!--     width: 50%; -->
<!--     height: auto; -->
<!--     display: block; -->
<!--     margin: 0 auto; /* Center the image horizontally */ -->
<!--     object-fit: cover; /* Crop the image to cover container */ -->
<!--   } -->

<!--   .text { -->
<!--     margin-top: 10px; -->
<!--     text-align: center; /* Center text within the box */ -->
<!--     text-decoration: none; -->
<!--   } -->

<!--   .text h1 { -->
<!--     font-size: 24px; /* Larger text size */ -->
<!--     margin-top: 1px; -->
<!--     margin-bottom: 5px; /* Adjust margin between large and small text */ -->
<!--     color: black; -->
<!--   } -->
<!--   .text h1:hover { -->
<!--     color: #2068DC; /* Change text color to blue on hover */ -->
<!--   } -->

<!--   .text p { -->
<!--     font-size: 14px; -->
<!--     margin-top: 5px; -->
<!--     color: #666; -->
<!--   } -->
<!-- </style> -->
<!-- </head> -->
<!-- <body> -->
<!--   <a href="wss.qmd" class="container"> -->
<!--     <div class="content"> -->
<!--       <img src="img/msu-masters/wss.JPG" alt="Image"> <!-- Center the image within the box --> 
<!--       <div class="text"> -->
<!--         <h1>Wheat Stem Sawfly</h1> -->
<!--         <p>Understanding the importance of smooth brome in promoting a Braconid parasitoid that targets the Wheat Stem Sawfly.</p> -->
<!--       </div> -->
<!--     </div> -->
<!--   </a> -->
<!--   <a href="locoweed-report.qmd" class="container"> -->
<!--     <div class="content"> -->
<!--       <img src="img/msu-masters/loco-bee2.jpg" alt="Image"> <!-- Center the image within the box -->
<!--       <div class="text"> -->
<!--         <h1>Locoweed Volatiles</h1> -->
<!--         <p>Comparison of volatile organic compounds profiles of <em>Oxytropis</em> plants that contained the commensal endophyte <em>Alternaria oxytropis</em> and those that did not.</p> -->
<!--       </div> -->
<!--     </div> -->
<!--   </a> -->
<!--   <a href="toadflax_report.qmd" class="container"> -->
<!--     <div class="content"> -->
<!--       <img src="img/msu-masters/toadflax.jpeg" alt="Image"> <!-- Center the image within the box --> 
<!--       <div class="text"> -->
<!--         <h1>Toadflax</h1> -->
<!--         <p></p> -->
<!--       </div> -->
<!--     </div> -->
<!--   </a> -->
<!-- </body> -->
<!-- </html> -->

```