### 2024-03-20 qmd/aboutme.qmd

---
title: "Jackson Strand"
about:
  id: me
  template: trestles
  links:
    - icon: linkedin
      text: LinkedIn
      href: https://linkedin.com/in/jacksonstrand
    - icon: instagram
      text: Instagram
      href: https://www.instagram.com/jstrand894/
    - icon: github
      text: GitHub
      href: https://github.com/jstrand894 
  image: ../img/aboutme/deadhorse.jpeg

sidebar: false
---

:::{#me}

I am currently a Master's graduate student at Montana State University studying entomology. I live in Bozeman, MT with my wife Grace.

```{=html}
<div  style="margin: 30px; text-align: center;">
  <a class="btn btn-primary" href="cv.qmd" role="button" style="padding: 15px 30px;">View my CV</a>
</div>
```

![Me and Grace at Glacier National Park, 2023](../img/aboutme/iceberg.JPG)

:::

<!-- ```{=html} -->
<!-- <!DOCTYPE html> -->
<!-- <html lang="en"> -->
<!-- <head> -->
<!--   <meta charset="UTF-8"> -->
<!--   <meta name="viewport" content="width=device-width, initial-scale=1.0"> -->
<!--   <title>Image Gallery</title> -->
<!--   <style> -->
<!--     .gallery { -->
<!--       display: grid; -->
<!--       grid-template-columns: repeat(auto-fill, minmax(200px, 1fr)); -->
<!--       grid-gap: 10px; -->
<!--     } -->
<!--     .gallery img { -->
<!--       width: 100%; -->
<!--       height: auto; -->
<!--       cursor: pointer; -->
<!--       transition: transform 0.3s ease; -->
<!--     } -->
<!--     .gallery img:hover { -->
<!--       transform: scale(1.05); /* Slightly increase size on hover */ -->
<!--     } -->
<!--     .expanded-container { -->
<!--       position: fixed; -->
<!--       top: 0; -->
<!--       left: 0; -->
<!--       width: 100%; -->
<!--       height: 100%; -->
<!--       background-color: rgba(0, 0, 0, 0.8); /* Semi-transparent background */ -->
<!--       display: none; /* Initially hidden */ -->
<!--       justify-content: center; -->
<!--       align-items: center; -->
<!--       z-index: 9999; -->
<!--     } -->

<!--     .expanded { -->
<!--       max-width: 80%; -->
<!--       max-height: 80%; -->
<!--       cursor: pointer; -->
<!--     } -->
<!--   </style> -->
<!-- </head> -->
<!-- <body> -->

<!-- <div class="gallery"> -->
<!--   <img src="../img/bugs/blueflowers.JPG" alt="Image 1" onclick="expandImage(this)"> -->
<!--   <img src="../img/bugs/brownmoth.JPG" alt="Image 2" onclick="expandImage(this)"> -->
<!--   <img src="../img/bugs/butterfly-purpleflower.JPG" alt="Image 3" onclick="expandImage(this)"> -->
<!-- </div> -->

<!-- <div class="expanded-container" onclick="closeExpanded()"> -->
<!--   <img class="expanded" id="expanded-image" onclick="navigateGallery(event)"> -->
<!-- </div> -->

<!-- <script> -->
<!--   var currentImageIndex = 0; -->
<!--   var galleryImages = document.querySelectorAll('.gallery img'); -->

<!--   function expandImage(img) { -->
<!--     currentImageIndex = Array.from(galleryImages).indexOf(img); -->
<!--     document.getElementById('expanded-image').src = img.src; -->
<!--     document.querySelector('.expanded-container').style.display = 'flex'; -->
<!--   } -->

<!--   function closeExpanded() { -->
<!--     document.querySelector('.expanded-container').style.display = 'none'; -->
<!--   } -->

<!--   function navigateGallery(event) { -->
<!--     // Calculate the click position relative to the expanded image -->
<!--     var clickX = event.pageX - event.target.getBoundingClientRect().left; -->
<!--     var imageWidth = event.target.offsetWidth; -->

<!--     // Determine if the click was on the left or right side of the image -->
<!--     if (clickX < imageWidth / 2) { -->
<!--       // Clicked on the left side: navigate to the previous image -->
<!--       currentImageIndex = (currentImageIndex - 1 + galleryImages.length) % galleryImages.length; -->
<!--     } else { -->
<!--       // Clicked on the right side: navigate to the next image -->
<!--       currentImageIndex = (currentImageIndex + 1) % galleryImages.length; -->
<!--     } -->

<!--     // Update the expanded image source -->
<!--     document.getElementById('expanded-image').src = galleryImages[currentImageIndex].src; -->
<!--   } -->
<!-- </script> -->

<!-- </body> -->
<!-- </html> -->

<!-- ``` -->