### 2024-03-20 qmd/cv.qmd

---
layout: page
title: "Cirriculum vitae"
execute:
  freeze: true  # never re-render during project render
engine: knitr 
resources:
  - "../Aboutme/jackson_strand_updated_cv.pdf"
cv:
  pdf: "../Aboutme/jackson_strand_updated_cv.pdf"
---

```{css echo=FALSE}
.embed-container {
    position: relative;
    padding-bottom: 129%;
    height: 0;
    overflow: hidden;
    max-width: 100%;
}
.embed-container iframe,
.embed-container object,
.embed-container embed {
    position: absolute;
    top: 0;
    left: 0;
    width: 100%;
    height: 100%;
}
```

```{=html}
<div class="embed-container">
  <iframe src="`r rmarkdown::metadata$cv$pdf`#toolbar=0" style="border: 0.5px;width: 100%;"></iframe>
</div>
```

```{=html}
<!DOCTYPE html>
<html lang="en">
<head>
<meta charset="UTF-8">
<meta name="viewport" content="width=device-width, initial-scale=1.0">
<title>Download File Button</title>
<style>
.container {
  display: flex;
  justify-content: center;
  align-items: center;
  height: 10vh;
}
/* Button styling */
.button {
  background-color: #2068DC; /* Header of website color (blue) */
  border: none;
  color: white;
  padding: 15px 32px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 16px;
  cursor: pointer;
  border-radius: 8px;
}
.button:hover {
  background-color: #3F91FB; /* Sky blue */
}
</style>
</head>
<body>

<div class = "container">
  <!-- Button to download the file -->
  <a href="../Aboutme/jackson_strand_updated_cv.pdf" download="Jackson Strand CV">
    <button class="button">Download</button>
  </a>
</div>

</body>
</html>
```