### 2024-03-20 qmd/projects/bballtool.qmd

---
title: "NBA Roster Analysis Tool"

editor_options: 
  chunk_output_type: console
---

From 2017 to 2019, I built my own NBA roster analysis tool for analyzing roster construction, free agency, and per-game metrics. The information is out of date, but all of the functionality still works.

```{=html}
<!DOCTYPE html>
<html lang="en">
<head>
<meta charset="UTF-8">
<meta name="viewport" content="width=device-width, initial-scale=1.0">
<title>NBA Roster Analysis Tool Preview</title>
</head>
<body>

<h2>Excel Preview</h2>

<!-- Replace 'path_to_your_excel_file.xlsx' with the actual path to your Excel file -->
<iframe src="https://1drv.ms/x/s!AiC9jtYNi0Ovhy_ySBYSU-xuO1CQ?e=ohHHit" width="100%" height="500px" frameborder="0">This browser does not support PDFs. Please download the PDF to view it: <a href="https://1drv.ms/x/s!AiC9jtYNi0Ovhy_ySBYSU-xuO1CQ?e=ohHHit</a></iframe>

</body>
</html>
```

Or download the file here:

```{=html}
<!DOCTYPE html>
<html lang="en">
<head>
<meta charset="UTF-8">
<meta name="viewport" content="width=device-width, initial-scale=1.0">
<title>Download File Button</title>
<style>
/* Button styling */
.button {
  background-color: #4CAF50; /* Green */
  border: none;
  color: white;
  padding: 15px 32px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 16px;
  margin: 4px 2px;
  cursor: pointer;
  border-radius: 8px;
}
.button:hover {
  background-color: #45a049; /* Darker green */
}
</style>
</head>
<body>

<!-- Button to download the file -->
<a href="/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/Personal/bballref.xlsx " download="Jackson Strand NBA Roster Analysis Tool.xlsx">
  <button class="button">Download File</button>
</a>

</body>
</html>

```