### 2024-03-25 qmd/aboutme.qmd

---
title: "Jackson Strand"
about:
  id: me
  template: trestles
  links:
    - icon: linkedin
      text: LinkedIn
      href: https://linkedin.com/in/jacksonstrand
    - icon: instagram
      text: Instagram
      href: https://www.instagram.com/jstrand894/
    - icon: github
      text: GitHub
      href: https://github.com/jstrand894 
  image: ../img/aboutme/deadhorse.jpeg

sidebar: false
---

:::{#me}

I am currently a Master's graduate student at Montana State University studying entomology. I live in Bozeman, MT with my wife Grace.

```{=html}
<div  style="margin: 30px; text-align: center;">
  <a class="btn btn-primary" href="cv.qmd" role="button" style="padding: 15px 30px;">View my CV</a>
</div>
```
:::

## Education  
<div class = "box">
**BA Biology**, 2013-2017 - *University of Minnesota - Twin Cities*    
    - Minor: Marine Biology

**MSc Entomology**, 2021-2024 - *Montana State University*
</div>

<!-- ![Me and Grace at Glacier National Park, 2023](../img/aboutme/iceberg.JPG)   -->