### 2024-03-26 index.qmd
updated index page

---
about: 
  template: jolla
  id: about-block
  image: "img/aboutme/crescent.jpg"
  links:
    - icon: linkedin
      text: LinkedIn
      href: https://linkedin.com/in/jacksonstrand
    - icon: envelope-fill
      text: Email
      href: "mailto:jstrand894@gmail.com"
    - icon: github
      text: GitHub
      href: https://github.com/jstrand894/
    - icon: instagram
      text: Instagram
      href: https://www.instagram.com/jstrand894/

toc: true
---
::: {#about-block}
<span style="font-size: 35px; font-family: 'Georgia', sans-serif;">**Jackson R. Strand**</span>

:::

## About me
I am currently a Master's graduate student at Montana State University studying entomology. I live in Bozeman, MT with my wife Grace.

**My Research Interests**  
I am broadly interested in how insects impact the lives of humans and how we can better utilize them to positively impact modern issues such as crop yield, climate change, and waste management.

**Montana State University - MSc Entomology**  
My research focuses on understanding how the ecological relationship between the wheat stem sawfly (*Cephus cinctus*) and two native biocontrol parasitoids (*Bracon cephi* and *Bracon lissogaster*) are impacted by smooth brome (*Bromis inermis*). My project includes field sites in southwest, central, and northern Montana, as well as volatile organic compound analysis of greenhouse-grown smooth brome and winter wheat.

<!-- pictures below about me -->
```{=html}
<!DOCTYPE html>
<html lang="en">
<head>
  <meta charset="UTF-8">
  <meta name="viewport" content="width=device-width, initial-scale=1.0">
  <title>Image Gallery</title>
  <style>
    .gallery-container {
      display: flex;
      justify-content: center;
      width: 100%;
    }
    .gallery {
      display: grid;
      grid-template-columns: repeat(auto-fill, minmax(150px, 1fr));
      grid-gap: 10px;
      max-width: 1200px;
      width: 100%;
    }
    .gallery img {
      width: 100%;
      object-fit: cover; /* Ensure images cover the entire space */
      cursor: pointer;
      transition: transform 0.3s ease;
    }
    .gallery img:hover {
      transform: scale(1.05); /* Slightly increase size on hover */
    }
    .expanded-container {
      position: fixed;
      top: 0;
      left: 0;
      width: 100%;
      height: 100%;
      background-color: rgba(0, 0, 0, 0.8); /* Semi-transparent background */
      display: none; /* Initially hidden */
      justify-content: center;
      align-items: center;
      z-index: 9999;
    }
    .expanded {
      max-width: 80%;
      max-height: 80%;
      cursor: pointer;
    }
    p.gallery-caption {
      text-align: center;
      font-style: italic;
      color: #8B8989;
    }
  </style>
</head>
<body>

<div class="gallery">
  <img src="img/index/wss.JPG" alt="Image 1" onclick="expandImage(this)">
  <img src="img/index/pinecreek.jpeg" alt="Image 1" onclick="expandImage(this)">
  <img src="img/index/loco-bee.JPG" alt="Image 1" onclick="expandImage(this)">
  <img src="img/index/mushroom1.jpeg" alt="Image 1" onclick="expandImage(this)">
  <img src="img/index/brownmoth.JPG" alt="Image 1" onclick="expandImage(this)">
  <img src="img/index/glacierfall.jpeg" alt="Image 1" onclick="expandImage(this)">
</div>
<div class="expanded-container" onclick="closeExpanded()">
  <img class="expanded" id="expanded-image" onclick="navigateGallery(event)">
</div>

<script>
  var currentImageIndex = 0;
  var galleryImages = document.querySelectorAll('.gallery img');

  function expandImage(img) {
    currentImageIndex = Array.from(galleryImages).indexOf(img);
    document.getElementById('expanded-image').src = img.src;
    document.querySelector('.expanded-container').style.display = 'flex';
  }

  function closeExpanded() {
    document.querySelector('.expanded-container').style.display = 'none';
  }

  function navigateGallery(event) {
    // Calculate the click position relative to the expanded image
    var clickX = event.pageX - event.target.getBoundingClientRect().left;
    var imageWidth = event.target.offsetWidth;

    // Determine if the click was on the left or right side of the image
    if (clickX < imageWidth / 2) {
      // Clicked on the left side: navigate to the previous image
      currentImageIndex = (currentImageIndex - 1 + galleryImages.length) % galleryImages.length;
    } else {
      // Clicked on the right side: navigate to the next image
      currentImageIndex = (currentImageIndex + 1) % galleryImages.length;
    }

    // Update the expanded image source
    document.getElementById('expanded-image').src = galleryImages[currentImageIndex].src;
  }
</script>

</body>
</html>

```

```{=html}
<div  style="margin: 30px; text-align: center;">
  <a class="btn btn-primary" href="qmd/cv.qmd" role="button" style="padding: 15px 30px;">View my CV</a>
</div>
```

<!-- ![Me and Grace at Glacier National Park, 2023](img/aboutme/iceberg.JPG)   -->

## Education

<div class="box">
<div style="text-align: justify;">
  <div style="float: left;"><strong>BS Biology</strong>
  <em>- University of Minnesota</em>
  </div>
  <div style="float: right;">2013-2017
  </div>
</div>
<br>

- Minor: Marine Biology

<div style="text-align: justify;">
  <div style="float: left;"><strong>MSc Entomology</strong>
  <em>- Montana State University</em>
  </div>
  <div style="float: right;">2021-2024
  </div>
</div>
<br>

- Committee: Dr. David Weaver (co-chair), Dr. Tracy Sterling (co-chair), Dr. Robert Peterson
- Graduate Certificate in Applied Statistics

</div>

## Research Experience
<div class="box">
<div style="text-align: justify;">
  <div style="float: left;"><strong>Montana State University</strong>
  </div>
  <div style="float: right;">2021-Present
  </div>
</div>
<br>

</div>

## Teaching Experience
<div class="box">
<div style="text-align: justify;">
  <div style="float: left;"><strong>TA for graduate/undergraduate level integrated pest management course</strong>
  </div>
  <div style="float: right;">Fall 2023
  </div>
</div>
<br>

  - Communicated weekly reading schedule and discussions to students.
  - Partook in online discussion on topics from class.
  - Was the immediate/primary contact for students.

</div>

## Employment History
<div class="box">
<!-- Pace Analytical Work -->
<div style="text-align: justify;">
  <div style="float: left;"><strong>
    <a href="https://www.pacelabs.com" target="_blank" style="color: inherit; text-decoration: none;">Pace Analytical - Onsite at 3M
    </a></strong> - <em>Regulatory IT Analyst</em>
  </div>
  <div style="float: right;">2017 - 2021
  </div>
</div>
<br>
(Maplewood, MN)

- Provided digital support between 3M employees and the 3M Chemical Data Management System (CDMS).  
- Integrated, tested, and supported user and data interface between CDMS and SAP.  
- Generated, designed, and communicated unique, detailed reports of material information using SQL, Microsoft Excel, and SAP.

<!-- Mosquito Control -->
<div style="text-align: justify;">
  <div style="float: left;"><strong>
    <a href="https://mmcd.org" target="_blank" style="color: inherit; text-decoration: none;">Metropolitan Mosquito Control District
    </a></strong> - <em>Seasonal Mosquito Technician</em>
  </div>
  <div style="float: right;">2017
  </div>
</div>
<br>
(Oakdale, MN)

- Worked on a team to collect and analyze mosquito larvae samples from wetland areas across the St. Paul metropolitan area.  
- Independently maintained low adult mosquito populations in Oakdale, MN by surveying and treating both adults and larvae.

<!-- Horse Barn -->
<div style="text-align: justify;">
  <div style="float: left;"><strong>
      <a href="https://vetmed.umn.edu/equine" target="_blank"
      style="color: inherit; text-decoration: none;">Leatherdale Equine Center
      </a>
    </strong> - 
      <em>Barn Maintenance Crew
      </em>
  </div>
  <div style="float: right;">2015-2017
  </div>
</div>
<br>
(St. Paul, MN)

- Maintained healthy conditions for equine patients at the University of Minnesota Horse Veterinary Center and Large Animal Hospital.  
- Trained and managed new members of the crew.

<!-- Soccer reffing -->
<div style="text-align: justify;">
  <div style="float: left;"><strong>Minnesota Youth Soccer Association</strong> - <em>Youth Soccer Referee</em>
  </div>
  <div style="float: right;">2010 - 2021
  </div>
</div>
<br>
</div>

## Honors & Awards
<!-- Entomology 2023 talk win -->
<div class="box">
<div style="text-align: justify;">
  <div style="float: left;"><strong>President's Prize - 2nd Place</strong>
  </div>
  <div style="float: right;">2023
  </div>
</div>
<br>

- Entomology 2023, National Harbor, MD  
- Entomological Society of America

<div style="text-align: justify;">
  <div style="float: left;"><strong>Student Scholar Travel Award</strong> ($500)
  </div>
  <div style="float: right;">2023
  </div>
</div>
<br>

  - College of Agriculture, Montana State University

<div style="text-align: justify;">
  <div style="float: left;"><strong>1st Place Speaker</strong>
  </div>
  <div style="float: right;">2023
  </div>
</div>
<br>

- 2023 LRES GSO Student Research Colloquium  
- LRES Graduate Student Organization, Montana State University

<div style="text-align: justify;">
  <div style="float: left;"><strong>Montana Grain Foundation Scholarship</strong> ($2,500)
  </div>
  <div style="float: right;">2022
  </div>
</div>
<br>

- Montana Grain Growers Association

<div style="text-align: justify;">
  <div style="float: left;"><strong>Professional Advancement Grant</strong> ($500)
  </div>
  <div style="float: right;">2021
  </div>
</div>
<br>

- The Graduate School, Montana State University
</div>

## Seminars & Presentations  
<div class="box">
**Paper Presentations**

1. Strand, J., Weaver, D.K. 2024. *How smooth brome can be beneficial for
parasitoids of the wheat stem sawfly.* March Madness – Montana Wheat and
Barley Committee. Montana State University, Bozeman, MT.

2. Sing, S.E., Ward, S.M., Peterson, R.K.D., Strand, J.R., Friesenhahn, E., Hofland,
M., Hubbard, C.E., Randall, C.B., Weaver. D.K. 2024. *Biological Control of Hybrid Toadflax: Insights from Chemical Ecology.* Montana State University, Bozeman, MT.

3. Strand, J., D.K. Weaver, T. Sterling, R.K.D. Peterson. 2023. *Using organic volatile compounds to assess the relationship between wheat stem sawfly, parasitoids, and smooth brome.* Session: Grad Competition, Plant-Insect Ecosystems: Chemical Ecology 2. <u>2nd Place</u>. Annual Meeting of the Entomological Society of America, National Harbor, MD.

4. Strand, J., D.K. Weaver, T. Sterling, R.K.D. Peterson. 2023. *The importance of smooth brome and parasitoids in wheat stem sawfly biocontrol.* <u>1st Place</u>. 2023 Land
Resources and Environmental Sciences Graduate Student Organization
Research Colloquium. Montana State University, Bozeman, MT.

5. Strand, J., Weaver, D.K., Sterling, T., Peterson, R.K.D. 2022. *Using organic
volatile compounds to assess the relationship between wheat stem sawfly, parasitoids, and smooth brome.* Session: Grad Competition, Plant-Insect Ecosystems: Chemical Ecology 1. Annual Meeting of the Entomological Society of America, Vancouver, BC, Canada.

6. Strand, J.R., Weaver, D.K., Sterling, T., Peterson, R.K.D. 2022. *Importance of parasitoids and smooth brome for integrated pest management of wheat stem sawfly.* 2022 Land Resources and Environmental Sciences Graduate Student Organization Research Colloquium, Montana State University, Bozeman, MT.

7. Strand, J.R., Sterling, T., Weaver, D.K., 2022. *Using organic volatiles to understand the impact of Alternaria oxytropis on locoweed physiology: E+/E-.* Multistate Research Project Meeting, W2193, "Poisonous Plants: Impact, Ecology, and Management." Montana State University, Bozeman, MT.

8. Strand, J.R., Weaver, D.K., Sterling, T., Peterson, R.K.D. 2021. *Importance of parasitoids and smooth brome for integrated pest management of wheat stem sawfly.* Session: Plant Insect-Ecosystems: IPM – General. Annual Meeting of the Entomological Society of America, Denver, CO.

**Workshops**

1. *Understanding the Wheat Stem Sawfly*. 2022 Integrated Pest Management
Workshop. Montana State University, Bozeman, MT.
</div>

## Service & Memberships
<div class="box">
**Society Memberships**
<!-- GSO co-chair -->
<div style="text-align: justify;">
  <div style="float: left;">LRES Graduate Student Organization (Co-chair)
  </div>
  <div style="float: right;">2023 - Present
  </div>
</div>
<br>

<!-- GSO Tresurer-->
<div style="text-align: justify;">
  <div style="float: left;">LRES Graduate Student Organization (Treasurer)
  </div>
  <div style="float: right;">2021 - 2023
  </div>
</div>
<br>

<!-- entomological society of america -->
<div style="text-align: justify;">
  <div style="float: left;">Entomological Society of America
  </div>
  <div style="float: right;">2021 - Present
  </div>
</div>
<br>

<!-- MGGA -->
<div style="text-align: justify;">
  <div style="float: left;">Montana Grain Growers Association
  </div>
  <div style="float: right;">2021 - Present
  </div>
</div>
<br>

**Volunteer**
<div style="text-align: justify;">
  <div style="float: left;">Southwest Montana Mountain Bike Association
  </div>
  <div style="float: right;">2021 - Present
  </div>
</div>
<br>

<div style="text-align: justify;">
  <div style="float: left;">Gallatin Watershed Council
  </div>
  <div style="float: right;">2021 - Present
  </div>
</div>
<br>

<div style="text-align: justify;">
  <div style="float: left;">Gallatin Valley Land Trust
  </div>
  <div style="float: right;">2021 - Present
  </div>
</div>
<br>

<div style="text-align: justify;">
  <div style="float: left;">Lake Elmo Regional Park Invasive Species Removal
  </div>
  <div style="float: right;">2021 - Present
  </div>
</div>
<br>