### 2024-03-28 qmd/helpful.qmd
added gso page

---
title: "Helpful Resources & Links"
---

## R Studio  
- Guide to using R made by [Jared Adam](https://www.linkedin.com/in/jared-adam-24b107241/){style="text-decoration: none;"}: 
<a href="https://jared-adam-psu.github.io/R-Tutorial-for-Tooker-Lab/" target="_blank">One code to rule them all (┛ಠ_ಠ)┛彡O</a>

## Quarto Website  
- YouTube video on how to make a R Studio Quarto Website: [How to create a web site using Quarto, RStudio, and GitHub Pages](https://youtu.be/YN75YXaLFGM?si=RyCG23IMBwFUKsO8)  
- Gallery showing different options for formatting a Quarto website: [Quarto Gallery](https://quarto.org/docs/gallery/)

## GitHub  
- A guide to using GitHub created by [Callum Arnold](https://www.linkedin.com/in/arnold-c/){style="text-decoration: none;"}: <a href="https://psu-git.callumarnold.com" target="_blank">Introduction to Git and GitHub</a>