### 2024-04-05 qmd/projects.qmd
updated locoweed field report

---
title: "Projects"
editor_options: 
  chunk_output_type: console
toc: true
---

# Montana State University

<!-- wss -->

<a href="projects/msu/wss.qmd" style="text-decoration:none;"> [Master's Project]{style="font-size:2em;"} </a> [*wheat stem sawfly and native parasitoids*]{style="fonts-size:1em;"}\

::: box
Understanding the ecological importance of smooth brome (*Bromus inermis*) in biological control of the wheat stem sawfly (*Cephus cinctus*) using native parasitoids. <br>\
*PI: [Dr. David Weaver](https://landresources.montana.edu/directory/faculty/1524725/david-weaver){style="text-decoration:none;"}*
:::

<br> 
<!-- locoweed --> 
<a href="projects/msu/locoweed.qmd" style="text-decoration:none;"> [Locoweed Volatiles]{style="font-size:2em;"} </a> [*volatile profile differences in Oxytropis sericea*]{style="fonts-size:1em;"}\

::: box
Measuring the differences in volatile organic compound (VOC) profiles between *Oxytropis sericea* specimens containing the commensal endophyte *Alternaira oxytropos*, and those without.\
<br> *PI: [Dr. Tracy Sterling](http://weedeco.msu.montana.edu/faculty/tracy_sterling/){style="text-decoration:none;"}*
:::

<br> 
<!-- toadflax --> 
<a href="projects/msu/toadflax.qmd" style="text-decoration:none;"> [Toadflax]{style="font-size:2em;"} </a> [*VOC data analysis*]{style="fonts-size:1em;"}\

::: box
Analyzing the difference in VOC profiles between two toadflax species and hybrids. Project in joint with the US Forest Service. <br>\
*PI: [Dr. David Weaver](https://landresources.montana.edu/directory/faculty/1524725/david-weaver){style="text-decoration: none;"} & [Dr. Sharlene Sing (USFS)](https://www.fs.usda.gov/research/about/people/ssing){style="text-decoration:none;"}*
:::

# Other

<a href="projects/gso.qmd" style="text-decoration:none;"> [LRES Graduate Student Organization]{style="font-size:2em;"} </a> \

::: box
Served as the treasurer from 2022-2023, then as co-chair from 2023-2024. Built and managed HTML emails, a website, and organized a research symposium.
:::

<br>

<a href="projects/bballtool.qmd" style="text-decoration:none;"> [NBA Roster Analysis Tool]{style="font-size:2em;"} </a> [basketball roster analysis]{style="fonts-size:1em;"}\

::: box
From 2017 to 2019, I built my own NBA roster analysis tool for analyzing roster construction, free agency, and per-game metrics. The information is out of date, but all of the functionality still works.
:::

<!-- ```{=html} -->
<!-- <!DOCTYPE html> -->
<!-- <html lang="en"> -->
<!-- <head> -->
<!-- <meta charset="UTF-8"> -->
<!-- <meta name="viewport" content="width=device-width, initial-scale=1.0"> -->
<!-- <title>Box with Rounded Edges</title> -->
<!-- <style> -->
<!--   body { -->
<!--     margin: 0; /* Remove default margin */ -->
<!--     padding: 0; /* Remove default padding */ -->
<!--   } -->

<!--   .container { -->
<!--     width: calc(100% - 20px); /* Adjusted width to fit the screen and account for margins */ -->
<!--     max-width: 600px; /* Limiting maximum width */ -->
<!--     height: auto; /* Set container height */ -->
<!--     border: 0.5px solid #ccc; /* Light grey color */ -->
<!--     border-radius: 8px; /* Less rounded edges */ -->
<!--     overflow: hidden; /* Hide overflow content */ -->
<!--     margin-bottom: 20px; /* Add margin to create space between rows */ -->
<!--     margin-right: auto; /* Push containers to the left */ -->
<!--     margin-left: 10px; /* Adjust margin between the left side of the screen */ -->
<!--     display: flex; /* Use flexbox for layout */ -->
<!--     align-items: center; /* Center items vertically */ -->
<!--     text-decoration: none; /* Remove underline */ -->
<!--   } -->

<!--   .container:last-child { -->
<!--     margin-bottom: 0; /* Remove margin from the last box */ -->
<!--   } -->

<!--   .container:hover { -->
<!--     box-shadow: 0 10px 20px rgba(0, 0, 0, 0.1); /* Increased shadow on hover */ -->
<!--   } -->

<!--   .image-container { -->
<!--     padding: 20px; -->
<!--     box-sizing: border-box; -->
<!--   } -->

<!--   .image-container img { -->
<!--     width: 80px; /* Adjusted width for smaller image */ -->
<!--     height: auto; -->
<!--     display: block; -->
<!--     border-radius: 8px; /* Rounded corners for image */ -->
<!--   } -->

<!--   .content { -->
<!--     padding: 20px; -->
<!--     box-sizing: border-box; -->
<!--     flex: 1; /* Take remaining space */ -->
<!--   } -->

<!--   .text h1 { -->
<!--     font-size: 22px; /* Adjusted text size */ -->
<!--     margin-top: 0; -->
<!--     margin-bottom: 5px; /* Adjust margin between large and small text */ -->
<!--     color: black; -->
<!--   } -->
<!--   .text h1:hover { -->
<!--     color: #2068DC; /* Change text color to blue on hover */ -->
<!--   } -->

<!--   .text p { -->
<!--     font-size: 16px; -->
<!--     margin-top: 5px; -->
<!--     color: #666; -->
<!--     text-align: justify; /* Align text */ -->
<!--   } -->

<!--   @media (max-width: 768px) { -->
<!--     .container { -->
<!--       width: calc(100% - 20px); /* Adjusted width to fit the screen */ -->
<!--     } -->
<!--   } -->
<!-- </style> -->
<!-- </head> -->
<!-- <body> -->
<!--   <a href="projects/msu.qmd" class="container"> -->
<!--     <div class="image-container"> -->
<!--       <img src="../img/projects-logos/msu-logo.jpg" alt="Image"> -->
<!--     </div> -->
<!--     <div class="content"> -->
<!--       <div class="text"> -->
<!--         <h1>Montana State University</h1> -->
<!--         <p>2021-2024</p> -->
<!--       </div> -->
<!--     </div> -->
<!--   </a> -->
<!--   <a href="projects/bballtool.qmd" class="container"> -->
<!--     <div class="image-container"> -->
<!--       <img src="../img/projects-logos/bballref-logo-circle.jpg" alt="Image"> -->
<!--     </div> -->
<!--     <div class="content"> -->
<!--       <div class="text"> -->
<!--         <h1>NBA Roster Analysis Tool</h1> -->
<!--         <p>2017-2019</p> -->
<!--       </div> -->
<!--     </div> -->
<!--   </a> -->
<!-- </body> -->
<!-- </html> -->
<!-- ``` -->
<!-- ```{=html} -->
<!-- <head> -->
<!--   <meta charset="UTF-8"> -->
<!--   <meta name="viewport" content="width=device-width, initial-scale=1.0"> -->
<!--   <title>Box with Rounded Edges</title> -->
<!--   <style> -->
<!--     body { -->
<!--       margin: 0; /* Remove default margin */ -->
<!--       padding: 0; /* Remove default padding */ -->
<!--     } -->

<!--     .container { -->
<!--       width: 100%; /* Adjusted width to fit the screen */ -->
<!--       max-width: 600px; /* Limiting maximum width */ -->
<!--       height: auto; /* Set container height */ -->
<!--       border: 0.5px solid #ccc; /* Light grey color */ -->
<!--       border-radius: 8px; /* Less rounded edges */ -->
<!--       overflow: hidden; /* Hide overflow content */ -->
<!--       margin-bottom: 20px; /* Add margin to create space between rows */ -->
<!--       display: flex; /* Use flexbox for layout */ -->
<!--       align-items: center; /* Center items vertically */ -->
<!--       text-decoration: none; /* Remove underline */ -->
<!--     } -->

<!--     .container:last-child { -->
<!--       margin-bottom: 0; /* Remove margin from the last box */ -->
<!--     } -->

<!--     .container:hover { -->
<!--       box-shadow: 0 10px 20px rgba(0, 0, 0, 0.1); /* Increased shadow on hover */ -->
<!--     } -->

<!--     .image-container { -->
<!--       padding: 20px; -->
<!--       box-sizing: border-box; -->
<!--     } -->

<!--     .image-container img { -->
<!--       max-width: 100%; /* Ensures the image does not exceed its container */ -->
<!--       height: auto; -->
<!--       display: block; -->
<!--       border-radius: 8px; /* Rounded corners for image */ -->
<!--     } -->

<!--     .content { -->
<!--       padding: 20px; -->
<!--       box-sizing: border-box; -->
<!--       flex: 1; /* Take remaining space */ -->
<!--     } -->

<!--     .text h1 { -->
<!--       font-size: 22px; /* Adjusted text size */ -->
<!--       margin-top: 0; -->
<!--       margin-bottom: 5px; /* Adjust margin between large and small text */ -->
<!--       color: black; -->
<!--     } -->

<!--     .text h1:hover { -->
<!--       color: #2068DC; /* Change text color to blue on hover */ -->
<!--     } -->

<!--     .text p { -->
<!--       font-size: 16px; -->
<!--       margin-top: 5px; -->
<!--       color: #666; -->
<!--       text-align: justify; /* Align text */ -->
<!--     } -->

<!--     @media (max-width: 768px) { -->
<!--       .container { -->
<!--         margin-right: auto; /* Reset margin-right */ -->
<!--         margin-left: auto; /* Reset margin-left */ -->
<!--         flex-direction: column; /* Stack items vertically on smaller screens */ -->
<!--       } -->

<!--       .content { -->
<!--         padding: 10px; /* Reduce padding on smaller screens */ -->
<!--       } -->
<!--     } -->
<!--   </style> -->
<!-- </head> -->
<!-- ``` -->
<!-- ```{r, echo = F, message = F, results = 'hide'} -->
<!-- library(pandoc) -->
<!-- library(knitr) -->
<!-- library(rmarkdown) -->
<!-- getwd() -->
<!-- rmarkdown::render(input = "litreview.qmd",  -->
<!--                   run_pandoc = TRUE, -->
<!--                   output_format = "word_document") -->

<!-- ``` -->