### 2024-05-13 qmd/helpful.qmd

---
title: "Helpful Resources & Links"
---

Below I have put many of the links that helped me.

## R Studio

-   Guide to using R made by [Jared Adam](https://www.linkedin.com/in/jared-adam-24b107241/){style="text-decoration: none;"}: <a href="https://jared-adam-psu.github.io/R-Tutorial-for-Tooker-Lab/" target="_blank">One code to rule them all (┛ಠ_ಠ)┛彡O</a>
-   Helpful free tutorial for learning R Studio through Codecademy: <a href="https://www.codecademy.com/learn/learn-r?utm_source=pepperjam&utm_medium=affiliate&utm_term=159404&clickId=4716939938&pj_creativeid=8-12462&pj_publisherid=159404" target="_blank">Learn R with Codecademy</a>
-   Learn how to use R Markdown to create pdf and word outputs with your code: <a href="https://rmarkdown.rstudio.com/lesson-1.html" target="_blank">R Markdown Documentation</a>

## Quarto Website

-   YouTube video on how to make a R Studio Quarto Website: [How to create a web site using Quarto, RStudio, and GitHub Pages](https://youtu.be/YN75YXaLFGM?si=RyCG23IMBwFUKsO8)

```{=html}
<!DOCTYPE html>
<html lang="en">
<head>
    <meta charset="UTF-8">
    <meta name="viewport" content="width=device-width, initial-scale=1.0">
    <title>How to create a web site using Quarto, RStudio, and GitHub Pages</title>
</head>
<body>
    <!-- <h2></h2> -->
    <!-- Replace 'VIDEO_ID' with the actual ID of your YouTube video -->
    <iframe width="560" height="315" src="https://www.youtube.com/embed/YN75YXaLFGM" frameborder="0" allowfullscreen></iframe>
</body>
</html>
```
-   Gallery showing different options for formatting a Quarto website: [Quarto Gallery](https://quarto.org/docs/gallery/)
-   Quarto documentation showing how to do a lot of the basics including website navigation, website search, and website tools: <a href="https://quarto.org/docs/websites/" target="_blank">Quarto - Creating a Website</a>

## GitHub

-   A guide to using GitHub created by [Callum Arnold](https://www.linkedin.com/in/arnold-c/){style="text-decoration: none;"}: <a href="https://psu-git.callumarnold.com" target="_blank">Introduction to Git and GitHub</a>

Watershed volunteering
https://youtu.be/65MAB0qRtAE?si=beO8jD5t5SOJzooF