### 2024-05-13 qmd/litreview.qmd

---
title: "Literature Review (Thesis)"
author: "Jackson Strand"
date: March 20, 2024
output: 
  html_document:
    toc: true
    code_folding: hide
bibliography: proposal.bib
link-citations: true
biblio-style: apalike
cls: elsevier-with-titles.csl
# cls: council-of-science-editors-author-date.csl
editor_options: 
  chunk_output_type: inline
toc: true
---

<!-- Button to download word document version -->

```{=html}
<!DOCTYPE html>
<html lang="en">
<head>
<meta charset="UTF-8">
<meta name="viewport" content="width=device-width, initial-scale=1.0">
<title>Download File Button</title>
<style>
/* Button styling */
.button {
  background-color: #2068DC; /* Header of website color (blue) */
  border: none;
  color: white;
  padding: 15px 32px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 16px;
  margin: 4px 2px;
  cursor: pointer;
  border-radius: 8px;
}
.button:hover {
  background-color: #3F91FB; /* Sky blue */
}
</style>
</head>
<body>

<!-- Button to download the file -->
<a href="litreview.docx" download="Jackson Strand Proposal">
  <button class="button">Download as Word Document</button>
</a>

<script>
// Get the current date in YYYY-MM-DD format
function getCurrentDate() {
  const now = new Date();
  const year = now.getFullYear();
  const month = String(now.getMonth() + 1).padStart(2, '0');
  const day = String(now.getDate()).padStart(2, '0');
  return `${year}-${month}-${day}`;
}

// Update the download attribute of the link with the current date
document.addEventListener('DOMContentLoaded', function() {
  const downloadLink = document.getElementById('downloadLink');
  const currentDate = getCurrentDate();
  downloadLink.setAttribute('download', `Jackson Strand Proposal ${currentDate}`);
});
</script>

</body>
</html>
```
# Project Overview

## Motivation

The motivation for my research is to improve yield for wheat growers by reducing wheat stem sawfly (WSS) populations. If we can show that healthier, more robust smooth brome leads to increased parasitoid populations and thus, increased WSS parasitism and potentially decreased damage by WSS, we can better educate growers on how to increase yields. In terms of scientific merit, understanding these mechanisms is key in ongoing WSS research as it explains several how and why questions associated with the ecological relationship between smooth brome and WSS parasitism.

## Research Question

What role does smooth brome (*Bromus inermis*) play in the ecological relationship between wheat stem sawfly (*Cephus cinctus*) and wheat stem sawfly parasitoids (*Bracon cephi* and *Bracon lissogaster*) in dryland wheat production landscape within Montana?

## Hypothesis

Smooth brome acts as an ecological sink for WSS, while serving as a parasitoid source because WSS larvae experience high mortality within smooth brome stems, while any survivors experience high rates of parasitism as a second source of mortality. In addition, early season parasitoids are attracted to infested smooth brome because it produces volatiles similar to those produced by winter wheat infested with larval WSS.

## Prediction

WSS larvae will experience both high levels of mortality and similar levels of parasitism within smooth brome stems when compared to the survivability and parasitism of WSS larvae in the adjacent fields. Also, the blend of volatile compounds produced by WSS infested smooth brome will qualitatively match in composition and exceed the amounts produced by infested winter wheat.

## Research Approach

My research is composed of three separate studies. First, there is a field survey to assess the survivability of WSS larvae and associated parasitoid frequency within both winter wheat and adjacent smooth brome. Second, there are controlled field infestations which will assess the survivability of WSS larvae within smooth brome stems when controlled for other density driven factors. And third, there is a greenhouse study to measure the volatile compounds produced by winter wheat and smooth brome. Using the results of the volatile collection, we can assess the parasitoid response both chemically with electroantennography and behaviorally with a Y-tube olfactory bioassay.

## Importance of Topic

The wheat stem sawfly (WSS, Cephus cinctus Norton, Hymenoptera: Cephidae), is a serious problem for wheat producers throughout the Northern Great Plains of North America [@Fulbright2017]. In Montana alone, the economic damage caused by the pest was reported to be upwards of \$80 million annually between 2011-2016 [@Fulbright2017]. Current techniques in combatting the damage caused by the WSS use a variety of chemical, cultural, and biocontrol strategies with limited short-term effectiveness and long-term sustainability. Utilizing biological control via promoting braconid parasitoid habitat holds considerable value to both growers and agricultural observers. As we seek to increase agricultural yields, we must also consider the environmental impact that tactics such as pesticides and tillage have on the future success on wheat farming.

# Literature Review

## History and Distribution of Wheat Stem Sawfly

The wheat stem sawfly was first described in 1872 by Edward Norton based upon a specimen found in Colorado [@Norton1872]. In 1890 it was found in California and Montana [@Ainslie1920]. Native to North America [@Lesieur2016], the largest populations of WSS are currently found in the western Canadian prairies as well as Wyoming, Montana, Colorado, and North and South Dakota [@Wallace1966; @Weiss1992; @Cockrell2021].

## Host Plants of Wheat Stem Sawfly

Wheat stem sawfly was first found the hollow-stemmed wild grasses such as Agropyron and Elymus in Colorado in 1872 [@Ainslie1920]. Soon thereafter it was found in wheat in 1895 in Manitoba, Canada. First referred to as the “western-grass stem sawfly”, WSS was found with native Montana grasses including western wheatgrass (Pascopyrun smithii) and slender wheat grass (Elymus trachycaulus) [@Ainslie1920]. As settlers began moving west and small grain cereal crop production increased on the landscape, the native grasses that WSS foraged upon became quite rare [@Ainslie1929]. WSS was found damaging spring wheat for the first time in 1895 in Manitoba and Saskatchewan [@Wallace1966]. With the diminishment of its natural hosts, WSS has adapted to infest the exotic non-cultivated and cultivated grasses now present [@Criddle1923]. WSS infestation has been shown in many species native to Montana including western wheatgrass (Pascopyrun smithii), slender wheat grass (Elymus trachycaulus) and bluebunch wheatgrass, Pseudoroegneria spicata [@Cockrell2021].

Smooth brome (*Bromus inermis* L.) is a Eurasian cool-season rhizomatous grass introduced to the United States in 1884 from Hungary and Russia for soil conservation and use as livestock feed [@Salesman2011]. Throughout the early 19th and 20th century, it was used throughout much of the upper half of North America. It is a perennial grass, and more drought resistant than other exotic grasses [@Mapfumo2002]. Due to its sod forming nature and large root system, it is extensively used to promote soil retention along highways (USDA-NRCS 2010).

Smooth brome was originally thought of as potentially useful as a trap crop in order to combat the growing WSS problem with cultural control practices [@Farstad1945]. However, due to its propensity to form monocultures and outcompete many native grasses, smooth brome was labeled by many as invasive [@Carlson1985; @Willson2000; @Dillemuth2008]. Smooth brome populations play a considerable role in altering native herbivore population dynamics and movement on the prairies of North America [@Dillemuth2008].

## Biology of Wheat Stem Sawfly

### Adult

Adults are relatively small (8-13 mm) with a narrow elongated and slightly compressed black body, yellow legs, and three characteristic black bands across the abdomen. Females are larger than males and a visible long, dark, ovipositor. Both male and female adults are weak fliers, rarely flying long distances at one time [@Criddle1923]. They can be found on stems with their heads downward and legs aligned with their body, encircling the stem [@Ainslie1920; @Criddle1923]. WSS are univoltine, having only one generation per year. Males are haploid (9 chromosomes), while females are diploid – females who do not find a mate will produce male offspring, while those who do produce female progeny [@Holmes1978]. Fully mature adults hatch in the spring once it has sufficiently warmed and the moisture levels in the soil reach threshold levels [@Holmes1978; @mendoza2006]. In Montana, this period begins in mid- to late-May and ends early in July [@Morrill1996; @Weaver2005]. Their lifespan usually lasts around five to eight days, emerging from May until June. Males emerge before the females, ensuring that many of the early females to hatch will find a mate and deposit eggs that will be female [@Mackay2011]. Copulation lasts less than one minute with the male holding the back of the female with their abdomen curved down under the female ovipositor. Fertilized females lay eggs several hours after copulation [@Holmes1978].

### Oviposition

After locating a suitable host stem, female WSS oviposit an egg inside the host plant lumen. WSS larvae hatch inside the stem and feed on the stem’s parenchyma tissue [@Beres2011]. As the stem begins to senesce in the fall, the larvae travel to the base of the stem and makes a V-shaped groove [@Criddle1923] above a constructed protective hibernaculum where they will overwinter [@Holmes1960]. As a result of this deep groove, the stem weakens and topples easily. The resulting lodged stems are not only challenging for growers to harvest but also exhibit decreased kernel yield [@Beres2011].

### Egg

Newly laid eggs are crescent in shape, milky-white in color, and range between 1 mm to 1.25 mm in length and 0.33 to 0.42 mm wide [@Ainslie1920]. After oviposition, the egg lies freely within the lumen of the host stem or in a hollow created by the female WSS ovipositor [@Fulbright2017]. Laval development occurs rapidly with the recognizable shape becoming apparent at day three. By day five, the jaws turn brown as the eye spots begin to appear and darken. By day six or seven, the larva begins to hatch using a series of convulsive movements and emerges into the stem cavity, with full functionality achieved by the seventh day [@Ainslie1920].

### Larvae

After hatching and prior to feeding, the larvae are transparent and colorless [@Ainslie1920; @Criddle1923]. WSS larvae are distinguishable from other larvae feeding in wheat and grasses by their highly sclerotized head capsule and large body size [@Nelson1953]. As the larvae chew their way through the parenchymous and vascular tissues of the host stem, the larvae begin to develop a yellow-green color. As the larvae progress through the stem, they excrete frass; the digested plant tissue as well as metabolic breakdown products from the WSS [@Beres2011]. If multiple eggs are laid within the same stem, the largest (usually the first egg laid and hatched) will typically remain within a few weeks due to the larvae cannibalizing one another [@Criddle1923; @Seamans1944; @Buteler2015]. It remains unknown whether the consumption of other eggs and larvae is the result of intentional cannibalism or the indirect result of indiscriminate feeding [@Buteler2015]. As the host plant physiologically matures, and the larvae reaches its fifth instar, the insect is cued to travel to the base of stem and prepare for obligatory diapause [@Beres2011]. The descending movement of larvae and development of the hibernaculum are triggered by the host plant senescence. As the stem walls becoming increasingly transparent, they allow increased amounts of visible and infrared light into the stem [@Holmes1977; @Beres2011]. Once the larvae reach the base of the stem, they construct an interior V-shaped groove, making the stem weak enough to lodge due to gravity or wind [@Criddle1923]. The larvae then plug the remaining stub with frass to insulate the hibernaculum during diapause [@Holmes1960]. The larval stage of development can be as short as 30 days before entering diapause. Obligatory diapause lasts a minimum of 90 days [@Salt1947]. After the larvae are exposed to temperatures less than 10°C for 90 days, diapause is complete, and pupation can begin. Larvae can reenter diapause if temperatures approach 35°C [@Salt1947] or if conditions are abnormally dry [@Holmes1978] shortly after termination of diapause.

### Pupae

Once diapause is complete, pupal development begins. Pupation lasts for 21 days, beginning in early- to mid-May [@Criddle1923]. Initially, the pupa is milky-white in color [@Ainslie1920], gradually changing to bright white as the wings develop [@Fulbright2017]. Additionally, as development continues, characteristic brown and black eye spots appear on the surface of the pupae. Once pupation is complete, the adult sawfly will chew through either the frass plug or the side of the stub, and emerge ready for flight [@Ainslie1920; @Holmes1960].

## Damage Caused by Wheat Stem Sawfly

Wheat stem sawfly larvae travel throughout the stem of the host plant, feeding on the parenchymal tissue and chewing through the nodes. As a result, vascular translocation of water and nutrients within the plant is impaired [@Macedo2007] and causes a notable reduction yield and quality of the grain produced [@Seamans1944; @Holmes1977; @Morrill1992]. In addition, the weakening of the stem due to WSS causes lodging prior to harvest, exacerbated by larvae stem girdling. Stem lodging causes a reduction in stem harvest efficiently [@Ainslie1920], with yield lost due to unrecoverable lodged stem [@Beres2007]. Additionally, the reduction in stubble residues decreases snow retention which increases soil erosion and reduces soil moisture content [@Keren2015].

## Management of Wheat Stem Sawfly

### Chemical Control

Because WSS are stem mining pests and spend most of their lives within the host plant stem, the use of traditional pest control methods such as insecticides have proven to be ineffective. The only life stage of WSS that takes place outside of the stem, the adult, is not targetable via chemical treatment due to adult emergence and oviposition taking place over many weeks [@Knodel2010; @Beres2011]. WSS infestation chemical control methods were first tested as early as 1949 [@Wallace1962]. After using six different organochlorine and organophosphate insecticidal dust applied via a powder duster [@Munro1949] the authors found “unsatisfactory results that did not control” WSS populations or limit infestations. In 1963, it was reported that a 73% control of WSS infestation could be achieved via granular heptachlor applied in the furrow with the wheat seed [@Wallace1962]. However, the findings of these tests were not able to be replicable in all scenarios of WSS infestation [@Holmes1963]. Heptachlor required light sawfly infestation – restricted to the lower internodes – to be effective. When infestation rates were high, and oviposition occurred in the upper internodes, the researchers saw little to no control. Similar results were found when testing with foliar insecticides, findings showing little differences in their effect when compared to controls [@Blodgett1996]. While the Montana Department of Agriculture has authorized the use of systemic organophosphate insecticide phorate (Thimet®) for use controlling WSS infestation in winter and spring wheat cultivars, it’s use is not considered environmentally friendly due to high toxicity to mammals, birds, and fish. Additionally, a preharvest interval of 85 days complicates harvest scheduling (Ginsberg, 2003; Mahajan et al., 2006), while the high toxicity may have potential adverse effects on beneficial parasitoid populations [@Varella2016].

### Cultural Control

#### Host Plant Resistance

Mechanisms like antixenosis, antibiosis, and tolerance give a plant the ability to resist insect pests [@Horber1980; @Wilson1987]. Antixenosis allows the host plant to trigger resistance by inducing non-preference in herbivores when they attempt to feed or oviposit. Recent work done by Achhami et al. (2020) found that Barely cultivars that closely resemble wheat show the greatest frequency of WSS oviposition. These findings suggest that semiochemicals produced by these barley cultivars, such as (Z)-3-hexenyl acetate [@Piesik2008; @Buteler2009] and (Z)-β-ocimene [@Buteler2012], are likely similar to those produced by wheat [@Achhami2020]. Antibiosis causes the improper development of herbivores, impacting survivability and reproduction. Tolerance allows the host plant to withstand infestation and overcome injuries [@VanEmden1991; @VanZoeren2017]. Much of the research done on WSS resistant wheat cultivars focuses on solid stem trait and pith expression [@Platt1949; @Varella2016; @Varella2017]. Solid stem varieties are more resistant to infestation, reducing WSS oviposition and increasing egg mortality [@Holmes1960; @Varella2016]. The increased pith expression also disrupts larvae movement within the stem, leading to higher larval mortality due to reduced feeding as well as smaller and less reproductively successful adults [@Wallace1966]. In addition, greater pith expression increases stem strength, reducing losses due to lodged stems [@Delaney2010]. Additional factors including late planting and shorter crop duration, have been suggested as potential factors complicating antixenosis, leading to increased WSS oviposition and infestation [@Achhami2020]. The solid-stemmed trait has been developed in numerous spring and some winter wheat cultivars [@Beres2011]. While the use of solid-stemmed cultivars is most often recommended for WSS management [@Beres2011], hollow-stemmed cultivars were found to exhibit greater parasitism when compared to solid-stemmed varieties [@Holmes1963]. It was previously suggested by Rand et al. (2012) that the thicker pith found in solid-stemmed cultivars makes sensing WSS larvae vibrations [@Mankin2004] more difficult for female parasitoids. The inability to locate WSS hosts would then cause ineffective parasitism.

#### Tillage

Tillage is a potential control method for the WSS. Larvae are protected against both water loss and freezing temperatures by the soil surrounding the wheat stubble where the WSS overwinters. Tilling the soil exposes the wheat stubble to the soil surface, conditions which kill the hibernating larvae [@Callenback1944]. Tillage of the WSS has been shown to be effective in decreasing the pest’s populations in both spring and winter wheat cultivars. Upwards of 90% mortality has been shown in North Dakota after fall tillage [@Weiss1987]. Many factors influence the effectiveness of tilling, including adequate exposure of wheat stubs on the soil surface and appropriate timing [@Holmes1956; @Morrill1992; @Weiss1992]. While the variability in effectiveness exists, some growers do consider it a crucial management tool [@Blodgett1996; @Lesieur2016].

While effective in killing WSS larvae, there are negative impacts caused by this practice. Water loss, soil compaction, and erosion have all been shown to be dramatically increased by tilling activities [@Phillips1980; @Lal1991]. In addition, tilling practices negatively impact beneficial populations of parasitoid wasps that target the WSS. WSS parasitoids overwinter higher in the stem compared to at the base like WSS [@Morrill1998; @Runyon2002]. By displacing soil when tilling, the upper portion of the plant is buried under the soil, preventing the parasitoid wasps from emerging in the spring [@Runyon2002; @Weaver2004].

#### Plant Density and Spacing

Historically, it has been observed that wheat fields with wider row spacing (12 inches) have greater WSS cutting when compared to fields with reduced spacing (3 to 6 inches). Additionally, fields planted at greater seed density with solid stem cultivars exhibit lower WSS cutting compared to reduced seed density [@Luginbill1958]. The reason for this relationship between decreased density and increased sawfly infestation is due to the moisture availability and height of the wheat stems. At higher densities, each stem has access to less water, in turn leading to the stem to develop a thicker pith – more solid stemmed. As discussed above, there is a negative linear relationship between an increase in the solid-stem characteristic and infestation [@Kemp1934].

#### Crop Rotation

Continually planting wheat year after year creates ideal conditions for the WSS [@Callenbach1945]. Rotating wheat with more resistant wheat varieties as well as with oats, or soybeans (crops that are less or not susceptible to WSS infestation) is recommended to decrease the year after year WSS infestation rates [@Lin2014]. The notable downside to utilizing crop rotation in a matter that is effective is the economic downside. Alternate-year, summer fallow production systems are typical for wheat growing practices across much of the northern great plains of the United States [@Weiss1992; @Troeh1999].

#### Trap Crops

Trap cropping utilized non-economically important plantings to help reduce populations or block movement of pests into valuable crop areas [@Hokkanen1991]. Effective trap crops of the WSS should have appealing characteristics to the WSS to attract the pest to the trap planting, and away from the wheat [@Farstad1940]. Historically, planted trap strips were used by attracting sawflies and destroying the strip at the end of the season, eliminating the larvae [@Morrill1992]. However, utilizing trap plantings of highly attractive wheat varieties has been shown to hold limited viability, as growers are not fond of destroying planted grain at the end of the growing season [@Goosey1999].

Alternative utilization of WSS traps entail planting and/or maintaining strips of smooth brome grass (Bromus inermis Leyss.) or other WSS host grasses [@Farstad1940]. Smooth brome in particular holds potential as it is a poor host to WSS larvae [@Bhandari2020], yet WSS adults continue to oviposit the stems. This relationship of greater oviposition with increased mortality may lead to decreases in overall WSS populations.

#### Delayed Planting & Swathing

Wheat planted later in the growing season avoids much of the damage caused by the WSS as the wheat reaches stem elongation after many of the adult sawflies have already laid their eggs [@Jacobson1952]. The problem with delayed planting is that the practice is known to make less efficient use the earl season moisture in the soil [@Morrill1996]. Growers may need to weigh the tradeoffs in delayed planting to reduce losses in areas with a history of severe infestation. Swathing, also known as windrowing, is the practice of cutting the wheat to accelerate the drying process and decrease the losses caused by WSS cutting [@Holmes1965; @Carlson2019; @McCullough2020]. By cutting the wheat and letting it dry prior to sawfly cutting, swathing works by reducing losses due to lodging [@Ainslie1920].

#### Biological Control Methods

The use of biocontrol in limiting WSS populations has been well studied with variable results. There are currently three types of organisms that have been reported as biocontrol agents of the WSS in wheat fields: Soil fungal pathogens, the predatory beetle Phyllobaenus dubius, and two parasitic Hymenoptera.

#### Soil Fungal Pathogens

Fungal pathogens are found throughout both the plant and animal kingdoms. Several fungi have been shown to negatively impact WSS larvae infestation within wheat stems. While highly lethal to sawfly larvae, F. graminearum, F. acuminatum, F. equiseti, and F. avenaceum also cause crown rot in the host plant. The trade-off between WSS control and crown rot may limit Fusarium spp. usefulness to WSS management. [@Wenda2009]. Findings by Tangtrakulwanich et al. (2014) found that the fungi B. bassiana and M. brunneum effectively reduced damage caused by WSS. Both fungi kill insects by infecting via spore contact or by insect consumption of treated plant tissue [@Portman2018].

#### Predatory Beetle

Phyllobaenus dubius (Wolcott) has been observed emerging from wheat stubble and within dissected wheat stems in Montana [@Morrill2001]. Further research is required on this coleopteran to determine its prospects going forward as an effective WSS management tool.

#### Parasitoids of the Wheat Stem Sawfly

Two parasitoid wasps are known target the wheat stem sawfly in wheat fields. The congeneric parasitoids, Bracon cephi (Gahan) and B. lissogaster Muesebeck, are the most effective agents of WSS biological control [@Runyon2001]. B. cephi and B. lissogaster, as distinguished by Runyon et al. (2001), were initially limited in their ability to parasitize WSS larvae within wheat stems. The wheat stem sawfly, originally known as the western grass-stem sawfly, was originally only found within native grasses throughout North America, existing in alongside B. cephi and B. lissogaster [@Ainslie1920; Cárcamo et al., 2012). However, as agriculture increased and native grasses were replaced by large monoculture landscapes, the WSS moved into wheat [@Farstad1940; @Davis1955; Holmes1978] while the parasitoids were slower to adapt [@Morrill1998].

## Biology of *Bracon cephi* and *Bracon lissogaster*

### Adult

Both *B. cephi* and *B. lissogaster* are protelean host-specific ectoparasitoids with two generations per year (bivoltine) [@Nelson1953; @Somsen1956; @Holmes1963]. They are ectoparasitoids and idiobionts, living on the integument of their hosts and immediately paralyzing their hosts in location and stage it was when it was attacked [@Nelson1953; @Runyon2002; @Weaver2004]. Male parasitoids are approximately 3.5 mm in length, while females are 4.1 mm. They are dark red, with black antennae, eyes, ovipositor and ocelli [@Gahan1918]. B. lissogaster and B. cephi are distinguishable by their head color and unique metasoma surface texture [@Runyon2001]. Adults emerge after completing a pupal stage within the host stem by chewing a hole through the stem wall. In Montana, the first generations are present from late June to late July, while the second generations are active in August and September [@Somsen1956]. Adult female B. cephi and B. lissogaster will locate a host stem using a combination of olfactory and volatile cues. The wasps will then locate a host sawfly larva using acoustic cues [@Mankin2004] and use their ovipositor first paralyze, then deposit 1-4 eggs next to the paralyzed larva.

### Egg

*B. cephi* and *B. lissogaster* eggs are on average 0.86 mm in length, exhibiting an elongated shape and pale-yellow color. Females lay one to three eggs on the WSS larval integument or adjacent to the larva. The eggs develop for one to three days before hatching [@Nelson1953; @Somsen1956].

### Larva

Parasitoid larvae are 1mm in length, with a translucent hue that turns increasingly brown as the larva progresses through five instar stages [@Nelson1953]. The newly hatched larvae have no appendages, save for antennae, and they exhibit highly sclerotized mouthparts and well-developed mandibles. After hatching, the B. lissogaster larva attach themselves to the integument of the host larva and immediately begin feeding [@Somsen1956]. When feeding is complete, little remains of the host but the integument. Following feeding, the larva will spin a lightly brown colored cocoon over the course of several days and attach to the inside wall of the host plant stem. The larval stage takes an average of ten days between initial hatch and cocoon completion. Because there are two generations per year, the early summer offspring will hatch and emerge from the host stem. Offspring laid as a part of the second generation will overwinter within the cocoons and emerge in the spring. [@Nelson1953]

### WSS and Parasitoid Interactions with Smooth Brome

Bhandari (2020) has shown that oviposition occurs with similar frequency in wheat as it does in smooth brome, however, we see much higher rates of larval mortality within smooth brome stems. This has led researchers to hypothesize about the feasibility of using smooth brome as a trap cop [@mendoza2006; @Lesieur2016]. While feasible as a feed crop, as it holds considerably less economic value compared to wheat or other high value crops [@Criddle1922]. Smooth brome may be usable in trap strips along the edge of wheat fields [@Hokkanen1991]. While a poor host for the wheat stem sawfly, surviving immature larvae found within brome stems can be parasitized by B. cephi and B. lissogaster. This combination of factors may indicate that smooth brome can function as a WSS sink and a parasitoid source, with very few WSS emerging from smooth brome patches, while also supporting populations of parasitoids that can then parasitize WSS in adjacent wheat fields.

## Plant Semiochemicals

Plants interact with other organisms through a complex system of chemical signals known as Semiochemicals. Semiochemicals are compounds or mixtures of compounds that are released from one organism that evoke a physiological or behavior response in another organism [@Agelopoulos1999]. Semiochemicals play a large role in both insect-insect and plant-insect interactions and are essential in insects locating feedings sites, mating sites, egg-laying sites, and/or refugia [@Prokopy1984].

While it is known that these chemicals play an important role in communication between plants, insects, and other organisms, there is still debate about the function of many of them [@Reddy2004; @Voelckel2018].

### Semiochemicals in Plant-Insect Systems

Within plant-insect systems, plants respond to insect herbivory by synthesizing and releasing volatile compounds and providing important signals to biocontrol agents to localize potential host insects [@Sharma2019]. In addition, semiochemicals provide localized signatures that allow insects to detect food sources, find mates, and locate preferential oviposition sites [@Norin2007]. Plants typically produce semiochemicals when defending themselves from herbivorous arthropods \[Guerrieri1999\] or pathogen attacks [@Smart2014]. These responses can be grouped into two categories, direct and indirect responses. Direct responses are compounds that the plant produces that have a direct effect on the herbivore or pathogen [@Metcalf1998]. Indirect responses are the compounds produced that have no impact directly on the pest, but instead may attract parasitoids, predators, or other biological control agents [@Mayland2000; @Colazza2004].

# Limitations And Key Assumptions

## Field Study

i.  Time limits the number field sites that we can collect samples from throughout the summer.\

```{=html}
<!-- -->
```
a.  Study area limited to western Montana – three regions – Big Sandy, Moccasin, and Amsterdam.\

```{=html}
<!-- -->
```
ii. Our study will only take place over the course of 2-3 summers, thus, we are unable to make broader claims surrounding long term patterns of WSS and parasitoid interactions.\
    Green House Samples and Volatile Collection
iii. We will assume that plants grown in the green house will be accurate and reasonable biological model representation of plants present in the field.\
iv. Volatiles will be collected from a maximum of 12 plants a day due to constraints with the collection system.

# Contributions to Knowledge of Potential Outcomes

## Potential outcomes

### Volatiles

We hope to observe a discernable difference in the volatile quantities produced by both infested smooth brome and winter wheat.

# Citations