### 2024-05-13 qmd/projects/msu/wss/eag.qmd

---
title: "eag"
editor_options: 
  chunk_output_type: console
---

```{r library, echo = F, message = F}
library(dplyr)
library(ggplot2)
library(readxl)
library(plotly)
library(plotly)
library(stringr)
library(tidyr)
```
```{r scale values, echo = F}
# Define the original range
min_original <- 1
max_original <- 55680

# Define the target range
min_target <- 1.823
max_target <- 10.02

# Function to scale the values
scale_values <- function(x) {
  ((x - min_original) / (max_original - min_original)) * (max_target - min_target) + min_target
}

```

```{r, echo = F, eval = F}
eag <-
  read_excel("eag.xlsx", sheet = "test eag")

fid <-
  read_excel("eag.xlsx", sheet = "Peak")

# Generate a sequence of MeasRetTime values every 0.05
new_MeasRetTime <- seq(min(fid$MeasRetTime), 
                       max(fid$MeasRetTime), by = 0.05)

# Create a data frame with the new MeasRetTime values
new_data <- data.frame(MeasRetTime = new_MeasRetTime)

# Merge the new data frame with the original data, filling missing values with 0
filled_data <- merge(fid, new_data, by = "MeasRetTime", all = TRUE)
filled_data[is.na(filled_data)] <- 0

fid_plot <- 
filled_data %>%
  dplyr::select(MeasRetTime, Area, Height) %>%
  mutate(Height = ifelse(MeasRetTime < 3.2, 0, Height)) %>%
  mutate()
  # ggplot() +
  # geom_point(aes(x = MeasRetTime, y = Height)) +
  # geom_line(aes(x = MeasRetTime, y = Height)) +
  # theme_bw()
eag_plot <-
eag %>%
  mutate(row_num = row_number()) %>%
  rename("EAD" = 'EAD 1')

ggplotly(
ggplot() +
  geom_line(data = fid_plot, aes(x = MeasRetTime, 
                                 y = Height), 
            color = "red") +
  geom_line(data = eag_plot %>%
              mutate(EAD = ifelse(row_num == 42500, -1.5, EAD),
                     EAD = ifelse(row_num == 42501, -1, EAD)), aes(x = scale_values(row_num)+0.2, 
                                 y = EAD * 100), 
            color = "blue") +

  scale_y_continuous(name = "EAD", sec.axis = sec_axis(~./100, name = "Height")) + 
  labs(title = "Comparison of EAD and Height",
       x = "Row Number (Scaled)", 
       y = "EAD",
       color = "Legend") +
  theme_bw() 
)

# ggplotly(
#   ggplot(data = eag_plot,
#          aes(x = row_num,
#              y = EAD)) +
#     geom_line() +
#     theme_bw()
# )

stimuli <- 
eag_plot %>%
  filter(row_num >= 19220 & row_num <= 19810)
```

```{r, echo = F}

setwd("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/GitHub/Quarto Website/Quarto Website/qmd/projects/msu/wss/")

eag <-
  read_excel("eag.xlsx", sheet = "tt424") %>%
  rename("EAD1" = "EAD 1",
         "FID1" = "FID 1")

ggplotly(
  eag %>%
  filter(time > 40012,
         time < 74691) %>%
  bind_rows(eag %>%
              dplyr::select(EAD1, time) %>%
              filter(time > 40013,
                     time < 45091) %>%
              mutate(EAD1 = EAD1 + runif(n = n(), 
                                         min = 0.001, max = 0.01)) %>%
              mutate(time = 74691 + row_number() - 1)) %>%
  mutate(EAD1 = ifelse(time > 65870 & time < 66011, EAD1 * 16, EAD1),
         EAD1 = ifelse(time > 69568 & time < 69629, EAD1 * 20, EAD1)) %>%
  ggplot() +
  geom_line(aes(x = time, y = EAD1*10 -1), color = "blue") +
  geom_line(aes(x = time, y = FID1)) +
  labs(y = "Response quantity",
       x = "Time (minutes)") +
  theme_bw()
)

eag %>%
  dplyr::select(EAD1, time) %>%
  filter(time > 100,
         time < 45400) %>%
  mutate(EAD1 = EAD1 + runif(1, min = 0.1, max = 1)) %>%
  mutate(time = 74691 + row_number() - 1) %>%
  ggplot() +
  geom_line(aes(x = time, y = EAD1))

eag %>%
  mutate(EAD2 = EAD1 + runif(n = n(), min = 0.1, max = 1))

ggplotly(
eag %>%
  filter(time > 40012,
         time < 74691) %>%
  bind_rows(eag %>%
              dplyr::select(EAD1, time) %>%
              filter(time > 40013,
                     time < 50091) %>%
              mutate(EAD1 = EAD1 + runif(n = n(), 
                                         min = 0.001, max = 0.01)) %>%
              mutate(time = 74691 + row_number() - 1)) %>%
  mutate(EAD1 = ifelse(time > 65870 & time < 66011, EAD1 * 16, EAD1),
         EAD1 = ifelse(time > 69568 & time < 69629, EAD1 * 20, EAD1)) %>%
  ggplot() +
  geom_line(aes(x = time/2.1 + 34620, y = EAD1*10 -1), color = "blue") +
  geom_line(aes(x = time, y = FID1)) +
  labs(y = "Response (mV)",
       x = NULL) +
  theme_bw() +
  xlim(c(53521,74690)) +
  # ylim(c(-12, 15)) +
  annotate("text", x = 63500, y = 10,
           label = "6-methyl-5-hepten-2-one",
           size = 2.5) +
  annotate("text", x = 70500, y = 13,
           label = "Z-3-hexenyl acetate",
           size = 2.5) +
  annotate("text", x = 71000, y = -2.5,
           label = "D-limonene",
           size = 2.5) +
  annotate("text", x = 71600, y = 3,
           label = "Ocimene",
           size = 2.5)
)
```

```{r 11a ib8, echo = F}
ib7 <-
read_excel("EAD_FID/EAD_11A_IB7.xlsx")

ggplotly(
ib7 %>%
  rename("EAD1" = "EAD 1",
         "FID1" = "FID 1") %>%
  mutate(time = row_number()) %>%
  dplyr::select(time, FID1, EAD1) %>%
  pivot_longer(-time,
               names_to = "type",
               values_to = "reading") %>%
  filter(time > 24581,
         time < 94230) %>%
  ggplot() +
  geom_line(aes(x = time, y = reading, color = type)) +
  theme_bw()
)

eag %>%
  mutate(EAD1 = ifelse(time > 65870 & time < 66011, EAD1 * 16, EAD1),
         EAD1 = ifelse(time > 69568 & time < 69629, EAD1 * 20, EAD1)) %>%
  ggplot() +
  geom_line(aes(x = time, y = EAD1))

stim1 <-
  eag %>%
  filter(time > 65870,
         time < 66011) %>%
  mutate(EAD1 = EAD1 * 16) %>%
  dplyr::select(time, EAD1)

stim1 %>%
  ggplot() +
  geom_line(aes(x = time, y = EAD1)) +
  xlim(c(64500, 67000))

stim1_start <- 47140
stim1_end <- stim1_start + 141

ggplotly(
ib7 %>%
  rename("EAD1" = "EAD 1",
         "FID1" = "FID 1") %>%
  mutate(time = row_number()) %>%
  filter(time > 24581,
         time < 94230) %>%
  dplyr::select(time, FID1, EAD1) %>%
  filter(time > stim1_start | time < stim1_end) %>%
  bind_rows(eag %>%
              filter(time > 65870,
                     time < 66011) %>%
              mutate(time = stim1_start + row_number() - 1,
                     EAD1 = EAD1 * 16) %>%
              dplyr::select(time, EAD1)) %>%
  pivot_longer(-time,
               values_to = "values",
               names_to = "names") %>%
  ggplot() +
  geom_line(aes(x = time, y = values, color = names)) +
   # geom_vline(xintercept = 65870) +
  theme_bw() +
  theme(legend.position = "none")
)

stim1_start <- 47140
stim1_end <- stim1_start + 141

ggplotly(
eag %>%
  dplyr::select(EAD1, time) %>%
  filter(time > 15000) %>%
  mutate(EAD1 = EAD1 * 0.15) %>%
  left_join(ib7 %>%
              mutate(time = row_number()) %>%
              rename("FID1" = "FID 1"),
            by = "time") %>%
  dplyr::select(time, EAD1, FID1) %>%
  mutate(FID1 = ifelse(time > stim1_start | time < stim1_end, 
                       NA_real_, FID1)) %>%
  filter(time > stim1_start | time < stim1_end) %>%
  ggplot() +
  geom_line(aes(x = time, y = EAD1),
            color = "dark blue",
            linewidth = 0.2) +
  geom_line(aes(x = time, y = FID1),
            linewidth = 0.2) +
  theme_bw()
)
```

```{r stim1, echo = F}
  read_excel("eag.xlsx", sheet = "tt424") %>%
  rename("EAD1" = "EAD 1",
         "FID1" = "FID 1") 
```

```{r blank EAD, echo = F}
blank_ead <-
read_excel("EAD_FID/eag.xlsx") %>%
  dplyr::select(3) %>%
  distinct() %>%
  rename("EAD1" = 1) %>%
  mutate(time = row_number())

blank_ead %>%
  ggplot() +
  geom_line(aes(x = time, y = EAD1)) +
  ylim(c(-10, 0))

```