### 2024-05-13 qmd/projects/msu/wss/greenhouse_volatiles.qmd

---
title: "greenhouse volatiles"
editor_options: 
  chunk_output_type: console
---

```{r library, echo = F, message = F, warning = F}
library(dplyr)
library(readxl)
library(ggplot2)
library(kableExtra)

```

```{r data, echo = F}
meta.23 <-
read_excel("greenhouse_volatiles/BWW 2023 Results.xlsx",
           sheet = "meta.data")

meta.22 <-
read_excel("greenhouse_volatiles/BWW 2022 Results.xlsx",
           sheet = "meta.data")
```

```{r filter, echo = F, eval = F}
meta.23 %>% 
  dplyr::select(Sample, rep, sample.id,
                inf, p.type) %>%
  filter(p.type == "b",
         inf == "y") %>%
  mutate(Value = as.numeric(gsub("\\D", "", rep))) %>%
  filter(rep == "11A")

meta.22 %>%
  dplyr::select(sample.id, infested, type) %>%
  filter(sample.id == "1A7")
  filter(infested == "Y",
         type == "B")

```