### 2024-05-13 qmd/projects/msu/wss/ndvi/ndvi.qmd

---
title: "ndvi"
editor_options: 
  chunk_output_type: console
---

```{r library, echo = F, message = F}
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
library(tidyr)
library(lubridate)
```

```{r big sandy, echo = F}
# setwd("/Users/jacksonstrand/Library/CloudStorage/OneDrive-MontanaStateUniversity/GitHub/Quarto Website/Quarto Website/qmd/projects/msu/wss")

lochy <-
read_excel("ndvi/ndvi_2023.xlsx", sheet = "lochy") %>%
  filter(use == "yes") %>%
  mutate(date = ymd(date1)) %>%
  dplyr::select(date,wheat,brome) %>%
  pivot_longer(-date,
               names_to = "plant_type",
               values_to = "ndvi")

lochy %>%
  ggplot() +
  geom_point(aes(x = date, y = ndvi, color = plant_type),
             # alpha = 0.5,
             pch = 1,
             size = 0.75) +
  geom_line(aes(x = date, y = ndvi, color = plant_type),
            alpha = 0.5, 
            size = 0.3) +
  geom_smooth(aes(x = date, y = ndvi, color = plant_type),
              se = FALSE) +
  theme_bw() +
  labs(x = NULL,
       y = "NDVI",
       color = "Plant Type",
       title = "NDVI comparison between B. inermis and adjacent wheat field")

```

This data is not as good. Maybe need to find a new site to investigate. 
```{r moccasin, echo = F}
moccasin <-
read_excel("ndvi_2023.xlsx", sheet = "moccasin") %>%
  dplyr::select(3:6) %>%
  rename("date" = 1,
         "use" = 3) %>%
  filter(use %in% c("yes",NA)) %>%
  mutate(date = ymd(date)) %>%
  dplyr::select(date,wheat,brome) %>%
  pivot_longer(-date,
               names_to = "plant_type",
               values_to = "ndvi") %>%
  filter(!date %in% c("2023-05-07",
                      "2023-05-12"))

ggplotly(
moccasin %>%
  ggplot() +
  geom_point(aes(x = date, y = ndvi, color = plant_type),
             alpha = 0.5,
             color = "black") +
  geom_line(aes(x = date, y = ndvi, color = plant_type),
            alpha = 0.5) +
  geom_smooth(aes(x = date, y = ndvi, color = plant_type),
              se = FALSE) +
  theme_bw() +
  labs(x = NULL,
       y = "NDVI",
       color = "Plant Type",
       title = "NDVI comparison between B. inermis and adjacent wheat field")
)
```

```{r plot, echo = F, eval = F}
ggplotly(data %>%
  filter(!date %in% c("2023-03-31",
                      "2023-04-20",
                      "2023-05-10",
                      "2023-05-30",
                      "2023-06-09",
                      "2023-06-19",
                      "2023-07-14",
                      "2023-08-13")) %>%
  # mutate(NDVI = ifelse(type == "wheat", -NDVI, NDVI)) %>%
  ggplot(aes(x = date, y = NDVI, fill = type)) +
  geom_bar(stat = 'identity', 
           position = position_dodge()) +
  geom_line(stat = "smooth") +
  theme_bw()
)
```