### 2024-05-13 qmd/projects/msu/wss/weatherdata.qmd

---
title: "Weather Data MT"
author: "Jackson Strand"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, echo = F, message = F}
library(tidyverse)
library(glue)
library(lubridate)
library(ggplot2)
library(maps)
library(readxl)

```

```{r, echo = F, message = F}
inventory_url <- "https://www.ncei.noaa.gov/pub/data/ghcn/daily/ghcnd-inventory.txt"

inventory <- read_table(inventory_url,
           col_names = c("station", "lat", "lon", "variable", "start", "end"))
```

```{r weather data load, echo = F}
area <- c("Big Sandy", "Moccasin", "Bozeman")
lats <- c(48.177540, 47.109558, 45.678871)
longs <- c(-110.114060, -109.884338, -111.060815)

coordinates <- data.frame(area, lats, longs)

my_coordinates <- 
coordinates %>%
  mutate(lats_conv = lats * 2 * pi/360,
         longs_conv = longs * 2 * pi/360) %>%
  ### SELECT AREA HERE ###
  filter(area == "Big Sandy")
  ########################

top_3_stations <- 
  inventory %>%
    dplyr::select(station, lat, lon, start, end) %>%
    mutate(lat_r = lat *2 *pi/360,
           lon_r = lon *2 *pi/360,
           d = 1.609344 * 3963 * acos((sin(lat_r) * sin(my_coordinates$lats_conv)) + cos(lat_r) * cos(my_coordinates$lats_conv) * cos(my_coordinates$longs_conv - lon_r))) %>%
    filter(start < 1960 & end > 2020) %>%
    dplyr::select(-c(start, end)) %>%
    distinct() %>%
    arrange(d) %>%
    top_n(n = -3, d)

station_1 <- 
  top_3_stations %>%
  slice(1) %>%
  pull(station)

station_2 <-
  top_3_stations %>%
  slice(2) %>%
  pull(station)

station_3 <-
  top_3_stations %>%
  slice(3) %>%
  pull(station)

####
my_station <- station_1
####
station_daily <- glue("https://ncei.noaa.gov/pub/data/ghcn/daily/by_station/{my_station}.csv.gz")

local_weather <- 
read_csv(station_daily,
         col_names = c("station", "date", "variable", "value", "a", "b", "c", "d")) %>%
  # filter(station == station_2) %>%
  dplyr::select(station, date, variable, value) %>%
  pivot_wider(names_from = "variable", values_from = "value",
                values_fill = 0) %>%
  dplyr::select(date, TMAX, PRCP, SNOW) %>%
  mutate(date = ymd(date)) %>%
  filter(date > "1950-01-01") %>%
  filter(TMAX > -100)

local_weather <-
local_weather %>%
  mutate(year = lubridate::year(date))
```

```{r plot, echo = F, message = F}
# plot the weather
local_weather %>%
  group_by(year) %>%
  filter(!year == "2024") %>%
  reframe(average_prcp = mean(PRCP, na.rm =TRUE)) %>%
  filter(year < 2024,
         year > 1975) %>%
  ggplot() +
  geom_point(aes(x = year, y = average_prcp)) +
  geom_line(aes(x = year, y = average_prcp)) +
  geom_smooth(aes(x = year, y = average_prcp), 
              method = "loess", 
              se = TRUE) +
  scale_x_continuous(breaks = seq(min(local_weather$year),
                                  max(local_weather$year),
                                  by = 5)) +
  labs(x = NULL,
       title = my_coordinates$area) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

local_weather %>%
  group_by(year) %>%
  reframe(average_prcp = mean(PRCP, na.rm =TRUE)) %>%
  filter(year >= 2010,
         year < 2024) %>%
  ggplot() +
  geom_point(aes(x = year, y = average_prcp)) +
  geom_line(aes(x = year, y = average_prcp)) +
  geom_smooth(aes(x = year, y = average_prcp), 
              method = "loess", 
              se = TRUE) +
  scale_x_continuous(breaks = seq(min(local_weather$year),
                                  max(local_weather$year),
                                  by = 1)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.minor = element_blank()) +
  labs(x = NULL,
       y = "Average Precipitation (inches)",
       title = paste0(my_coordinates$area, ", MT", " Average Yearly Precipitation", collapse =""))
```

Let's run a linear model to see if the downward trend is significant.
```{r linear model, echo = F}
local_weather %>%
  filter(year >= 2010,
         year < 2024) %>%
  group_by(year) %>%
  reframe(average_prcp = mean(PRCP, na.rm =TRUE)) %>%
  lm(average_prcp ~ year, data = .) %>%
  summary()
```

```{r map, echo = F}
world_map <- map_data("world")
states <- map_data("state")
counties <- map_data("county") %>%
  filter(region == "montana")

ggplot() +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  geom_polygon(data = states, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", alpha = 0.5) +
  geom_polygon(data = counties, aes(x = long, y = lat, group = group),
               fill = NA, color = "black", alpha = 0.5) +
  geom_label(data = coordinates, aes(x = longs, y = lats), 
             color = "black", 
             label = area,
             size = 3,
             vjust = -0.3,
             hjust = -0.1,
             label.padding = unit(0.3, "lines")) +
  geom_point(data = coordinates, aes(x = longs, y = lats),
             color = "red") +
  coord_cartesian(xlim = c(-116, -104), ylim = c(44, 49)) +
  theme_bw()

```

```{r}

library(pacman)
rgee_environment_dir = "/opt/miniconda3/envs/rgee_py"

p_load(raster,
       rgdal,
       rgeos,
       gdalUtils,
       sp,
       sf,
       leaflet,
       mapview,
       caret)

p_load(rgee, geojsonio, remotes, reticulate, devtools, googledrive)
install_github("r-spatial/rgee")

library(remotes)
library(rgee)

# set python
reticulate::use_python(rgee_environment_dir, required=T)
rgee::ee_install_set_pyenv(
  py_path = rgee_environment_dir, # Change it for your own Python PATH
  py_env = "rgee_py" # Change it for your own Python ENV
)
Sys.setenv(RETICULATE_PYTHON = rgee_environment_dir)
Sys.setenv(EARTHENGINE_PYTHON = rgee_environment_dir)
```

Big Sandy Data
- Need data from 3 weather stations that are nearby lochy's property
- average the data to get a good representation 
- plot the past 30 years of data
- plot the past 7 years of data

```{r}
library(readxl)
lochy_weather <- 
read_excel("Lochy_weather.xlsx", sheet = "clean2") %>%
  mutate(date = ymd(date)) %>%
  dplyr::select(region, lat, long, date, year, prcp, tmax) %>%
  group_by(year) %>%
  reframe(prcp = sum(prcp * 0.3937 / 2)) +
  mutate(year = as.factor(year))

```

```{r Big Sandy Data, echo = F}
area <- c("Big Sandy", "Moccasin", "Bozeman")
lats <- c(48.177540, 47.098790, 45.678871)
longs <- c(-110.114060, -109.882962, -111.060815)

coordinates <- data.frame(area, lats, longs)

my_coordinates <- 
coordinates %>%
  mutate(lats_conv = lats * 2 * pi/360,
         longs_conv = longs * 2 * pi/360) %>%
  ### SELECT AREA HERE ###
  filter(area == "Big Sandy")
  ########################

top_3_stations <- 
  inventory %>%
    dplyr::select(station, lat, lon, start, end) %>%
    mutate(lat_r = lat *2 *pi/360,
           lon_r = lon *2 *pi/360,
           d = 1.609344 * 3963 * acos((sin(lat_r) * sin(my_coordinates$lats_conv)) + cos(lat_r) * cos(my_coordinates$lats_conv) * cos(my_coordinates$longs_conv - lon_r))) %>%
    filter(start < 1960 & end > 2020) %>%
    dplyr::select(-c(start, end)) %>%
    distinct() %>%
    arrange(d) %>%
    top_n(n = -3, d) %>%
    mutate(miles = d / 1.609344)

# station 1
station_1 <- 
  top_3_stations %>%
  slice(1) %>%
  pull(station)

station_1_data <-
  read_csv(glue("https://ncei.noaa.gov/pub/data/ghcn/daily/by_station/{station_1}.csv.gz")) %>%
           set_names(c("station", "date", "variable", 
                         "value", "a", "b", "c", "d")) %>%
  dplyr::select(station, date, variable, value)

# station 2
station_2 <-
  top_3_stations %>%
  slice(2) %>%
  pull(station)

station_2_data <-
  read_csv(glue("https://ncei.noaa.gov/pub/data/ghcn/daily/by_station/{station_2}.csv.gz")) %>%
           set_names(c("station", "date", "variable", 
                         "value", "a", "b", "c", "d")) %>%
  dplyr::select(station, date, variable, value)

# station 3
station_3 <-
  top_3_stations %>%
  slice(3) %>%
  pull(station)

station_3_data <-
  read_csv(glue("https://ncei.noaa.gov/pub/data/ghcn/daily/by_station/{station_3}.csv.gz")) %>%
           set_names(c("station", "date", "variable", 
                         "value", "a", "b", "c", "d")) %>%
  dplyr::select(station, date, variable, value)

bind_rows(station_1_data,
          station_2_data,
          station_3_data) %>%
  pivot_wider(names_from = "variable", values_from = "value",
                values_fill = 0) %>%
  dplyr::select(station, date, TMAX, PRCP, SNOW) %>%
  mutate(date = ymd(date)) %>%
  filter(date > "1950-01-01") %>%
  filter(TMAX > -100) %>%
  mutate(year = lubridate::year(date)) %>%
  group_by(station, year) %>%
  reframe(average_temp = mean(TMAX),
          average_prcp = mean(PRCP)) %>%
  filter(year < 2024,
         year > 2014) %>%
  left_join(top_3_stations, by = "station") %>%
  mutate(miles = as.factor(round(miles,1))) %>%
  ggplot() +
  geom_point(aes(x = year, y = average_prcp, color = miles)) +
  geom_line(aes(x = year, y = average_prcp, color = miles)) +
  geom_smooth(aes(x = year, y = average_prcp), 
              method = "loess", 
              se = TRUE) +
  scale_x_continuous(breaks = seq(min(local_weather$year),
                                  max(local_weather$year),
                                  by = 1)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.minor = element_blank()) +
  labs(y = "Average Precipitation (inches)",
       x = NULL)

```

```{r}
# ndvi_brome <-
  read_excel("ndvi_fixed.xlsx") %>%
  dplyr::select(NDVI, date) %>%
  mutate(date = ymd(date)) %>%
  mutate(type = "brome") %>%
  bind_rows(ndvi_wheat) %>%
  mutate(year = lubridate::year(date),
         month = lubridate::month(date),
         day = lubridate::day(date)) %>%
  mutate(year = as.factor(year)) %>%
  # group_by(year, month, type) %>%
  # reframe(date, NDVI = mean(NDVI)) %>%
  ungroup() %>%
  filter(year == "2021",
         # NDVI > 0.05,
         month > 3,
         month < 10) %>%
  ggplot() +
  geom_line(aes(x = date, y = NDVI, color = type)) +
  theme_bw()

read_excel("ndvi_brome_nn_2.xlsx") %>%
  dplyr::select(NDVI, date) %>%
  mutate(date = ymd(date),
         year = lubridate::year(date)) %>%
  filter(year == "2023") %>%
  ggplot() +
  geom_line(aes(x = date, y = NDVI))

ndvi_wheat <-
read_excel("ndvi_10years_nn.xlsx") %>%
  dplyr::select(NDVI, date) %>%
  mutate(date = ymd(date)) %>%
  mutate(type = "wheat")

  filter(month > 3,
         month < 10,
         NDVI > 0.05) %>%
  mutate(year = as.factor(year),
         month = as.factor(month)) %>%
  filter(year == "2023")

ndvi_data <-
read_excel("ndvi_fixed.xlsx") %>%
  mutate(type = "brome") %>%
  bind_rows(read_excel("wheat_fixed.xlsx") %>%
              mutate(type = "wheat")) %>%
  dplyr::select(NDVI, date, type) %>%
  mutate(date = ymd(date),
         year = lubridate::year(date),
         month = lubridate::month(date))

ndvi_data %>%
  mutate(subset = "original") %>%
  filter(year == "2020") %>%
  bind_rows(ndvi_data %>%
              filter(year == "2023") %>%
              filter(!(NDVI < 0.05 & month > 4 & month < 10),
                     !(NDVI > 0.15 & month > 7)) %>%
              mutate(subset = "new")) %>%
  filter(subset == "new") %>%
  ggplot() +
  geom_point(aes(x = date, y = NDVI, color = type), 
             alpha = 0.7,
             position = position_jitter(width = 0.001, height = 0.002)) +
  geom_smooth(aes(x = date, y = NDVI, color = type),
              alpha = 0.5) +
  theme_bw() 
```