### 2024-05-17 qmd/projects/msu/wss/convert.Rmd
updated website for last time today hopefully 7/17/24

---
title: "Convert"
author: "Jackson Strand"
date: "`r Sys.Date()`"
output: html_document
---

```{r convert2}
# create a word document version of this document
# first need to copy and paste all of the content from this 
# document into an .RMD file so that I can convert the .RMD
# into a word document
library(rmarkdown)
content <- readLines("paper_1.qmd", warn = FALSE)
writeLines(content, "paper1.rmd")
render("paper1.rmd", 
       output_format = "word_document",
       output_file = paste0("JacksonStrand_Paper 1.docx"))
render("paper1.rmd", output_format = "pdf_document",
       output_file = paste0("JacksonStrand_Paper 1.pdf"))

# move the files to the convert folder
file.copy(paste0("JacksonStrand_Paper 1.docx"), 
          paste0("convert/JacksonStrand_Paper 1.docx"), overwrite = TRUE)
file.remove("JacksonStrand_Paper 1.docx")

file.copy(paste0("JacksonStrand_Paper 1.pdf"), 
          paste0("convert/JacksonStrand_Paper 1.pdf"), overwrite = TRUE)
file.remove("JacksonStrand_Paper 1.pdf")
```